---
Author: David and Keri
title: '20170717'
output: html_document
---

Setting up. Function to read in multiple files, selecting:
Year, Student ID, DBN, Grade_Level, and Date
```{r setup}
knitr::opts_chunk$set(echo = TRUE, root.dir = '/data/nycdoe/')
options(tibble.width=Inf)

library(ggplot2)
library(readr)
library(tidyverse)
files <- Sys.glob('/data/nycdoe/June Biog/*.csv')

read_csv_with_year <- function(filename) {
  year <- as.numeric(gsub('-.*$', '', basename(filename)))
  df <- read_csv(filename, col_types = cols(student_id_scram = col_character(), grade_level = col_character()))
  df$year <- year
  df <- select(df, year, student_id_scram, dbn, grade_level, contains("date"))
}

```
Calling in Data
```{r using bios df}
setwd('/data/nycdoe/')
files <- Sys.glob('/data/nycdoe/June Biog/*.csv')
bios <- map_df(files, read_csv_with_year)
```
#Some Weird Stuff noting
```{r using bios df}
bios %>%
  filter(grade_level=='PK') %>%
  group_by(year) %>%
  count()
bios %>%
  filter(grade_level=='0K') %>%
  group_by(year) %>%
  count()
bios %>%
  filter(grade_level=='IN') %>%
  group_by(year) %>%
  count()
bios %>%
  filter(grade_level=='AD') %>%
  group_by(year) %>%
  count()
```

#Some Weird Stuff noting
```{r using bios df}
bios %>%
  group_by(student_id_scram) %>%
  count() %>%
  arrange(n) %>%
  head()
```
Shaping the Data Frame

```{r pressure, echo=TRUE}
bios <- bios %>%
  mutate(date = ifelse(!is.na(admit_date), admit_date, date)) %>%
  select(-contains("_date"))

#fix this later
bios <- bios %>% mutate(date= floor(date/10000))
```
Convertingg grade level to numeric
```{r, echo=TRUE}
bios <- bios %>%
  mutate(grade_level = ifelse(grade_level=='IN', '-2', grade_level)) %>%
  mutate(grade_level = ifelse(grade_level=='0K', '0', grade_level)) %>%
  mutate(grade_level = ifelse(grade_level=='PK', '-1', grade_level)) %>%
  mutate(grade_level = as.numeric(grade_level))

bios <- bios %>%
  filter(is.na(grade_level))
```
Getting date entered, date exited, elementary admit date, high school admit date
```{r, echo=TRUE}
bios_when_entered <-
  bios %>%
  group_by(student_id_scram) %>%
  mutate(date_entered = ifelse(date==min(date), 1, 0 ), date_exit = ifelse(year == max(year), 1, 0 ) , 
         elementary = ifelse(grade_level<6 & date== min(date),1,0))
bios_when_entered %>% head(10)
```
Getting high school admit date
```{r, echo=TRUE}
bios_highchool <-
  bios %>%
  filter(grade_level > 8) %>%
  group_by(student_id_scram) %>%
  mutate(high = ifelse(date == min(date), 1, 0)) %>%
  select(year, date, student_id_scram, high)
bios_highchool %>% head(10)
```
Getting middle achool admit date
```{r, echo=TRUE}
bios_middleschool <-
  bios %>%
  filter(grade_level > 5,grade_level< 9) %>%
  group_by(student_id_scram) %>%
  mutate(middle = ifelse(date == min(date),1,0)) %>%
  select(year, date, student_id_scram, middle)
bios_middleschool %>% head(10)

```
joining middle school and high school data on to the main datafram
```{r, echo=TRUE}

bios_all <- left_join(bios_when_entered, bios_middleschool, by = c("year","student_id_scram","date"))
bios_all <- left_join(bios_all, bios_highchool, by = c("year", "student_id_scram","date"))
bios_all %>% head(10)

```
Converting NA to 0 for a binary column
```{r, echo=TRUE}
bios_all <- bios_all %>%
  mutate(middle = ifelse(is.na(middle),0,1)) %>%
  mutate(high = ifelse(is.na(high),0,1))
bios_all %>% head(10)
#Example with a few random students
bios_all %>% filter(student_id_scram == 840002100)
bios_all %>% filter(student_id_scram == 644002000)
bios_all %>% filter(student_id_scram == 420002000) 

```
Looking for students who appear in system for 10+ years
```{r, echo=TRUE}
bios_all %>%
  group_by(student_id_scram) %>%
  count() %>%
  filter(n>=10) %>%
  arrange(student_id_scram) %>%
  head(50)
#Note: find students who appear for 10 years
#find students in low profile/high profile schools
```
Looking at 5 students who are in system for 11 years
```{r, echo=TRUE}
bios_all %>% filter(student_id_scram == 100002032)
bios_all %>% filter(student_id_scram == 100012130)
bios_all %>% filter(student_id_scram == 100012639) 
bios_all %>% filter(student_id_scram == 100012899)
bios_all %>% filter(student_id_scram == 10002611) 

```
Keri trying to spread data frame
```{r, echo=TRUE}
#spread_bios <-
#  bios %>% 
#  group_by(student_id_scram) %>% 
#  spread(year, grade_level)
```
 