---
title: "20170719-Plot Application Rates"
author: "Anandini, Thoa"
date: "7/17/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
#setwd('/data/nycdoe/')
hsaps_files <- Sys.glob('/data/nycdoe/HSAPS/*.csv')
june_biog_files <- Sys.glob('/data/nycdoe/June Biog/*.csv')
```

Here we read in HSAPS data from 2009-10 to 2013-14. We only choose columns that has `r1programcode*` which specifies the program that a student ranks and applies to in the first round.

```{r read-hsaps, include=FALSE}
# read hsaps data
read_apps <- function(filename) {
  print(filename)
  year <- as.numeric(gsub('-.*$', '', basename(filename)))
  df <- read_csv(filename, col_types = cols(student_id_scram=col_character()))
  df$year <- year
  df <- df %>% 
    select(year, student_id_scram, contains("r1programcode"))
}

apps_0910_1314 <- map_df(hsaps_files[5:9], read_apps)
```

We also read June Biog data in to have info about 8th grade in 2009-10 to 2013-14.

```{r read-junebiog, include=FALSE}
# read juneBiog data
read_students_hs <- function(filename) {
  year <- as.numeric(gsub('-.*$', '', basename(filename)))
  df <- read_csv(filename, col_types = cols(student_id_scram=col_character(),
                                            grade_level=col_character()))
  df$year <- year
  df <- df %>% 
    select(year, student_id_scram, grade_level, dbn)
}
students_hs_0910_1314 <- map_df(june_biog_files[5:9], read_students_hs)
```

## Sanity check:

Good news: students always fill out higher-ranked choices before lower-ranked ones. (That's why the following code results in nothing)

```{r sanity-check}
apps_0910_1314 %>%
  gather("choice", "program_applied", 3:14) %>%
  mutate(foo = is.na(program_applied)) %>%
  select(year, student_id_scram, choice, program_applied, foo) %>%
  arrange(year, student_id_scram) %>%
  group_by(year, student_id_scram) %>%
  summarize(bar = length(rle(foo)),
            baz = sum(lag(foo, default=0) != foo)) %>%
  filter(bar > 2 | baz > 1)
```

## Filter out applicants that are not 8th-graders.

### Method 1: select the first year the applicant appears in the data.
```{r method-1, include=FALSE}
(k8_applicants_0910_1314_method1 <-
  apps_0910_1314 %>%
  group_by(student_id_scram) %>%
  filter(year == min(year)))
```

### Method 2: verify applicants' grade using June Biog data.
```{r method-2, include=FALSE}
(students_grades <-
  inner_join(students_hs_0910_1314, apps_0910_1314, by=c("student_id_scram", "year")) %>%
  filter(grade_level == "08"))

(k8_applicants_0910_1314_method2 <-
  apps_0910_1314 %>%
  filter(student_id_scram %in% students_grades$student_id_scram))

k8_applicants_0910_1314_method2 %>%
  group_by(student_id_scram) %>%
  summarize(num_years = length(unique(year))) %>%
  arrange(desc(num_years)) %>%
  group_by(num_years) %>%
  count()
```

The two methods yield different results. The first method guarantees all student_id's have one single year entry. The second method (we think) guarantees applicants are current 8th-graders, but a fraction of them have multiple year entries (entered the application process, got matched, but then got stuck in 8th-grade?). The first method yields around 50,000 more rows than the second method. We think there's likelihood that some first-time applicants might be missing in June Biog data 8th grade because they are not in DOE in 8th grade.

## What percent of all 8th-grade applicants each school get in a year?

We need to calculate the total number of applicants per year, as follows. First table is from Method 1's data. Second table is from Method 2's data.

```{r total-up-applicants, echo=FALSE}
(total_applicants_per_year_method1 <-
  k8_applicants_0910_1314_method1 %>%
  filter(!is.na(r1programcode1)) %>% 
  group_by(year) %>% 
  summarize(total_applicants = length(unique(student_id_scram))))
(total_applicants_per_year_method2 <-
  k8_applicants_0910_1314_method2 %>%
  filter(!is.na(r1programcode1)) %>% 
  group_by(year) %>% 
  summarize(total_applicants = length(unique(student_id_scram))))
```

We first need to gather `r1programcode*` from wide to long, so each `student_id_scram` will have 12 rows of programcode (some rows will be NA if the student does not use all of the choices).

### If the program appears in one of the top 12 in 1st round.


```{r top-12, include=FALSE}

(top12_applications_method1 <- 
  k8_applicants_0910_1314_method1 %>%
  gather("choice", "program_applied", 3:14) %>%
  filter(!is.na(program_applied)) %>%
  group_by(year, program_applied) %>%
  summarize(num_applications = n()) %>%
  group_by(year) %>%
  mutate(pc_applications = num_applications/sum(num_applications)) %>%
  left_join(total_applicants_per_year_method1, by="year") %>%
  mutate(app_rate = num_applications/total_applicants) %>%
  mutate(pc_applications = as.numeric(format(round(pc_applications, digits=5), nsmall=5)),  
         app_rate = as.numeric(format(round(app_rate, digits=5), nsmall=5))) %>%
  arrange(year, desc(app_rate)))

(top12_applications_method2 <- 
  k8_applicants_0910_1314_method2 %>%
  gather("choice", "program_applied", 3:14) %>%
  filter(!is.na(program_applied)) %>%
  group_by(year, program_applied) %>%
  summarize(num_applications = n()) %>%
  group_by(year) %>%
  mutate(pc_applications = num_applications/sum(num_applications)) %>%
  left_join(total_applicants_per_year_method2, by="year") %>%
  mutate(app_rate = num_applications/total_applicants) %>%
  mutate(pc_applications = as.numeric(format(round(pc_applications, digits=5), nsmall=5)),  
         app_rate = as.numeric(format(round(app_rate, digits=5), nsmall=5))) %>%
  arrange(year, desc(app_rate)))
```
### If the program appears in one of the top 6 in 1st round.

```{r top-6, include=FALSE}
(top6_applications_method1 <- 
  k8_applicants_0910_1314_method1 %>%
  gather("choice", "program_applied", 3:8) %>%
  filter(!is.na(program_applied)) %>% 
  group_by(year, program_applied) %>%
  summarize(num_applications = sum(!is.na(program_applied))) %>%
  group_by(year) %>%
  mutate(pc_applications = num_applications/sum(num_applications)) %>%
  left_join(total_applicants_per_year_method2, by="year") %>%
  mutate(app_rate = num_applications/total_applicants) %>%
  mutate(pc_applications = as.numeric(format(round(pc_applications, digits=5), nsmall=5)),  
         app_rate = as.numeric(format(round(app_rate, digits=5), nsmall=5))) %>%
  arrange(year, desc(app_rate)))

(top6_applications_method2 <- 
  k8_applicants_0910_1314_method2 %>%
  gather("choice", "program_applied", 3:8) %>%
  filter(!is.na(program_applied)) %>%
  group_by(year, program_applied) %>%
  summarize(num_applications = sum(!is.na(program_applied))) %>%
  group_by(year) %>%
  mutate(pc_applications = num_applications/sum(num_applications)) %>%
  left_join(total_applicants_per_year_method2, by="year") %>%
  mutate(app_rate = num_applications/total_applicants) %>%
  mutate(pc_applications = as.numeric(format(round(pc_applications, digits=5), nsmall=5)),  
         app_rate = as.numeric(format(round(app_rate, digits=5), nsmall=5))) %>%
  arrange(year, desc(app_rate)))
```

### If the program appears in one of the top 3 in 1st round.

```{r top-3, include=FALSE}
(top3_applications_method1 <- 
  k8_applicants_0910_1314_method1 %>%
  gather("choice", "program_applied", 3:5) %>%
  filter(!is.na(program_applied)) %>% 
  group_by(year, program_applied) %>%
  summarize(num_applications = sum(!is.na(program_applied))) %>%
  group_by(year) %>%
  mutate(pc_applications = num_applications/sum(num_applications)) %>%
  left_join(total_applicants_per_year_method2, by="year") %>%
  mutate(app_rate = num_applications/total_applicants) %>%
  mutate(pc_applications = as.numeric(format(round(pc_applications, digits=5), nsmall=5)),  
         app_rate = as.numeric(format(round(app_rate, digits=5), nsmall=5))) %>%
  arrange(year, desc(app_rate)))

(top3_applications_method2 <- 
  k8_applicants_0910_1314_method2 %>%
  gather("choice", "program_applied", 3:5) %>%
  group_by(year, program_applied) %>%
  summarize(num_applications = sum(!is.na(program_applied))) %>%
  group_by(year) %>%
  mutate(pc_applications = num_applications/sum(num_applications)) %>%
  left_join(total_applicants_per_year_method2, by="year") %>%
  mutate(app_rate = num_applications/total_applicants) %>%
  mutate(pc_applications = as.numeric(format(round(pc_applications, digits=5), nsmall=5)),  
         app_rate = as.numeric(format(round(app_rate, digits=5), nsmall=5))) %>%
  arrange(year, desc(app_rate)))
```

### If the program appears in one of the top 1 in 1st round.

```{r top-1, include=FALSE}
(top1_applications_method1 <- 
  k8_applicants_0910_1314_method1 %>%
  gather("choice", "program_applied", 3:3) %>%
  filter(!is.na(program_applied)) %>% 
  group_by(year, program_applied) %>%
  summarize(num_applications = sum(!is.na(program_applied))) %>%
  group_by(year) %>%
  mutate(pc_applications = num_applications/sum(num_applications)) %>%
  left_join(total_applicants_per_year_method2, by="year") %>%
  mutate(app_rate = num_applications/total_applicants) %>%
  mutate(pc_applications = as.numeric(format(round(pc_applications, digits=5), nsmall=5)),  
         app_rate = as.numeric(format(round(app_rate, digits=5), nsmall=5))) %>%
  arrange(year, desc(app_rate)))

(top1_applications_method2 <- 
  k8_applicants_0910_1314_method2 %>%
  gather("choice", "program_applied", 3:3) %>%
  group_by(year, program_applied) %>%
  summarize(num_applications = sum(!is.na(program_applied))) %>%
  group_by(year) %>%
  mutate(pc_applications = num_applications/sum(num_applications)) %>%
  left_join(total_applicants_per_year_method2, by="year") %>%
  mutate(app_rate = num_applications/total_applicants) %>%
  mutate(pc_applications = as.numeric(format(round(pc_applications, digits=5), nsmall=5)),  
         app_rate = as.numeric(format(round(app_rate, digits=5), nsmall=5))) %>%
  arrange(year, desc(app_rate)))
```


## Match program codes to school codes

### Read and clean up data:

We read in, merge and clean up the data that map programs (name & code) to schools (name & code).

* Note 1: from OpenData NYC, we are able to obtain data of three years: 2016, 2014-15, 2013-14. 
* Note 2: we read in two types of tables. One maps program codes to dbn's. The other maps dbn's to school names.


```{r, include=FALSE}
library(readr)
codes2016 <- read_csv(url('https://data.cityofnewyork.us/api/views/ge8j-uqbf/rows.csv?accessType=DOWNLOAD'))
codes201415 <- read_csv(url('https://data.cityofnewyork.us/api/views/by6m-6zpb/rows.csv?accessType=DOWNLOAD'))
codes201314 <- read_csv(url('https://data.cityofnewyork.us/api/views/i9pf-sj7c/rows.csv?accessType=DOWNLOAD'))

names2016 <- read_csv(url('https://data.cityofnewyork.us/api/views/7crd-d9xh/rows.csv?accessType=DOWNLOAD'))
names201415 <- read_csv(url('https://data.cityofnewyork.us/api/views/n3p6-zve2/rows.csv?accessType=DOWNLOAD'))
names201314 <- read_csv(url('https://data.cityofnewyork.us/api/views/u553-m549/rows.csv?accessType=DOWNLOAD'))

codes2016 %>% filter(!(dbn %in% names201314$dbn))
codes201415 %>% filter(!(dbn %in% names201415$dbn))  #identify the "insane" row and drop it.
codes201415 <-
  codes201415 %>%
  filter(dbn %in% names201415$dbn)
codes201314 %>% filter(!(DBN %in% names201314$DBN))

(schoolcodes2016 <-
  codes2016 %>%
  mutate(year = 2016) %>%
  full_join(names2016, by="dbn") %>%
  select(year, pgcode = code, pgname = program, dbn, school_name, interest, method)) %>%
  print(width = Inf)
(schoolcodes201415 <-
  codes201415 %>%
  mutate(year = 201415) %>%
  full_join(names201415, by="dbn") %>%
  select(year, pgcode = code, pgname = program, dbn, school_name, interest, method)) %>%
  print(width = Inf)
(schoolcodes201314 <-
  codes201314 %>%
  mutate(year = 201314) %>%
  full_join(names201314, by="DBN") %>%
  select(year,
         pgcode = CODE,
         pgname = PROGRAM,
         dbn = DBN,
         school_name = Printed_Name,
         interest = INTEREST,
         method = METHOD)) %>%
  print(width = Inf)

allschoolcodes <-
  rbind(schoolcodes201314,
        schoolcodes201415,
        schoolcodes2016) 

####### OLD CODE ##########################################
# allschoolcodes %>%
#   group_by(pgcode, dbn, year) %>%
#   count() %>%
#   filter(n != 1)
# allschoolcodes %>%
#   filter(dbn == "03M485")
# # there are 8 programs that have no code, and they all belong to year 2013-14, and they all belong to one of the 8 specialized school (already excluding LaGuardia M485)
# allschoolcodes %>% 
#   filter(is.na(pgcode))
# # we now try looking up these dbn's in the entire table, to see if they are present in other years or not.
# allschoolcodes %>%
#   filter(dbn %in% c("10X445",
#                     "31R605",
#                     "02M475",
#                     "05M692",
#                     "10X696",
#                     "13K430",
#                     "14K449",
#                     "28Q687") & year != 201314)
##############################################################
```

Then we do some sanity check, as follows:

```{r sanity check}
# this code helps us find out that programs offered at the 9 specialized schools do not have codes.
allschoolcodes %>% 
  filter(is.na(pgcode) | pgcode == "NO CODE")

# because there are no program codes there, we will drop these schools from our table.
# MANUAL WAY:
# allschoolcodes <-
#   allschoolcodes %>%
#   filter(!(dbn %in% c("03M485",
#                       "10X445",
#                       "31R605",
#                       "02M475",
#                       "05M692",
#                       "10X696",
#                       "13K430",
#                       "14K449",
#                       "28Q687")))
# SMARTER WAY:
allschoolcodes <-
  allschoolcodes %>%
  filter(!is.na(pgcode) & pgcode != "NO CODE")

# now every program appears only once per school in each year.
allschoolcodes %>%
  group_by(pgcode, dbn, year) %>%
  count() %>%
  filter(n != 1)

# even more precise: each program is unique to each school
allschoolcodes %>%
  group_by(pgcode) %>%
  summarize(num_schools = length(unique(dbn))) %>%
  filter(num_schools != 1)

# but one school may have multiple programs
allschoolcodes %>%
  group_by(dbn) %>%
  summarize(num_programs = length(unique(pgcode))) %>%
  filter(num_programs != 1)
```

```{r map programs to schools in app_rate tables, include=FALSE}
# because each program is unique to each school, we now collapse the table so that there's only 1 row per programcode
minimal_allschoolcodes <-
  allschoolcodes %>%
  mutate(year = ifelse(year == 2016,
                       yes = 201600,
                       no = year)) %>%
  group_by(pgcode) %>%
  filter(year == max(year))
  
# join the schools to the programcode's % applicants table. Since there are too many tables and graphs, for now, we will use data obtained from Method 1 above.
top12_apps_and_schools_method1 <-
  top12_applications_method1 %>%
  left_join(minimal_allschoolcodes, by=c("program_applied" = "pgcode")) %>%
  select(year = year.x,
         total_applicants,
         pgcode = program_applied,
         pgname,
         dbn,
         school_name,
         num_applications,
         app_rate,
         pc_applications)

# There are roughly 600 program codes appearing in 2009 to 2013 that do not match to any school dbn.
top12_apps_and_schools_method1 %>%
  filter(is.na(dbn)) %>%
  distinct(year)
```


## What are some of the most "popular" HS in each year?

Our initial definition of popularity is the highest percentages of applicants a school receives in a year.
Later on, we want to incorporate the number of seat availability and calculate the ratio of num_apps : seats_avail. The higher this ratio is, the more popular a school is.

Since there are too many tables and graphs, for now, we will use data obtained from Method 1 above.

Note: when we did the mapping of programs to schools, there were 574 program codes in 2009 to 2013 that do not match to any school dbn. 

The following code selects out the top 10 most popular school in EACH year: 
```{r popularity}
school_apprates <- 
  top12_apps_and_schools_method1 %>%
  filter(!is.na(dbn)) %>%
  group_by(year, dbn, school_name) %>%
  summarize(num_applications = sum(num_applications),
            total_applicants = mean(total_applicants),
            app_rate = num_applications/total_applicants) %>%
  arrange(year, desc(app_rate)) %>%
  group_by(year) %>%
  top_n(10)
```

## Plot % applicants to different programs

### Top 10 schools in each year
```{r plot-top-schools-per-year, echo=FALSE}
# top 10 schools in each year
yr <- 2009
school_apprates %>%
  filter(year == as.numeric(yr)) %>%
  arrange(app_rate) %>%
  ggplot(aes(x=app_rate,y=factor(school_name, levels = (school_name)))) +
  labs(x="% applicants",y=paste("Top 10 schools in", yr), title=yr) +
  geom_point()
yr <- 2010
school_apprates %>%
  filter(year == as.numeric(yr)) %>%
  arrange(app_rate) %>%
  ggplot(aes(x=app_rate,y=factor(school_name, levels = school_name))) +
  labs(x="% applicants",y=paste("Top 10 schools in", yr), title=yr) +
  geom_point()
yr <- 2011
school_apprates %>%
  filter(year == as.numeric(yr)) %>%
  arrange(app_rate) %>%
  ggplot(aes(x=app_rate,y=factor(school_name, levels = school_name))) +
  labs(x="% applicants",y=paste("Top 10 schools in", yr), title=yr) +
  geom_point()
yr <- 2012
school_apprates %>%
  filter(year == as.numeric(yr)) %>%
  arrange(app_rate) %>%
  ggplot(aes(x=app_rate,y=factor(school_name, levels = school_name))) +
  labs(x="% applicants",y=paste("Top 10 schools in", yr), title=yr) +
  geom_point()
yr <- 2013
school_apprates %>%
  filter(year == as.numeric(yr)) %>%
  arrange(app_rate) %>%
  ggplot(aes(x=app_rate,y=factor(school_name, levels = school_name))) +
  labs(x="% applicants",y=paste("Top 10 schools in", yr), title=yr) +
  geom_point()
```

### Comparing the % applicants of different schools

Schools here are ranked based on a metric we named "aggregate app rate" which is the ratio of:

* numerator: summing across all years the num_applicants each school receives each year
* denominator: summing across all years the total_applicants each year

```{r plot-top-compared-across-years, echo=FALSE}
temp <- 
  top12_apps_and_schools_method1 %>%
  filter(!is.na(dbn)) %>%
  group_by(year, dbn, school_name) %>%
  summarize(num_applications = sum(num_applications),
            total_applicants = mean(total_applicants)) %>%
  group_by(dbn, school_name) %>%
  summarize(general_app_rate = sum(num_applications)/sum(total_applicants)) %>%
  #mutate(general_app_rate = as.numeric(format(round(general_app_rate, digits=5), nsmall=5))) %>%
  arrange(desc(general_app_rate))

temp$rank <- rep(c(1:ceiling(nrow(temp)/10)),
                 length.out = nrow(temp),
                 each = 10)

top12_fooz_method1 <- 
  top12_apps_and_schools_method1 %>%
  filter(!is.na(dbn)) %>%
  group_by(year, dbn, school_name) %>%
  summarize(num_applications = sum(num_applications),
            total_applicants = mean(total_applicants),
            app_rate = num_applications/total_applicants)
top10_schools <-
  temp %>%
  filter(rank == 1) %>%
  select(dbn, school_name) %>%
  left_join(top12_fooz_method1, by="school_name") %>%
  mutate(group = "Top 10 schools")

middle10_schools <-
  temp %>%
  filter(rank == as.integer(mean(temp$rank))) %>%
  select(dbn, school_name) %>%
  left_join(top12_fooz_method1, by="school_name") %>%
  mutate(group = "Middle 10 schools")

top_vs_middle <-
  top10_schools %>%
  rbind(middle10_schools)

top_vs_middle %>%
  ggplot() +
  geom_line(aes(x = year,
                y = app_rate,
                color = dbn.x),
            size = 0.75) +
  labs(y = "% applicants", color = "Schools", title = "Comparing schools in different popularity groups") +
  facet_wrap(~group)
```

### Distribution of % applicants

* Top N: an applicant lists a program anywhere in the top N of their choice list.
* Method 1 selects the first year each applicant appears in the HSAPS data. It runs the risk of missing out applicants who might have applied for 9th grade in more than one year, e.g. because they were held back in 8th grade.
* Method 2 selects applicants that are listed in June Biog as 8th-grader in the year they applied. The risk is that June Biog data might repeat entries for some 8th-graders although they were not necessarily held back.

```{r, include=FALSE}
library(scales)
```

```{r plot-distribution-of-applicants, echo=FALSE}
require(ggplot2)
# base::prettyNum will display pretty numbers (i.e. 0.1 instead of 1e-01 or 100 instead of 1e+2) up to 1e-4, then switch to scientific numbers. The following function lowers the limit to less than 1e-5:
prettyNum0 <- function(x){ifelse(x > 2e-5, sprintf("%.5g", x), "0.00001")} 
# if not want pretty numbers, can use this to display exponents of 10: labels = trans_format("log10", math_format(10^.x));
# or use this to display scientific numbers: labels = trans_format("log10", function(x) 10^x)

###### METHOD 1's DATA #########
# TOP 12
#################################################################################################
top12_fooz_method1 %>%
  ggplot(aes(x=app_rate))+
  geom_histogram(bins = 100, fill = "dark blue") +
  scale_x_continuous(label = percent) + 
  theme(panel.grid.minor = element_blank()) +
  facet_wrap(~year) +
  labs(title = "Distribution of % applicants - TOP12 - method 1",
       #subtitle = "Applicant means s/o who puts the programcode anywhere in their 12 choices ",
       subtitle = "Method 1 selects the first year each applicant appears in the HSAPS data.",
       y = "# schools",
       x = "% applicants - x-scale is continous")

top12_fooz_method1 %>%
  ggplot(aes(x=app_rate))+
  geom_histogram(bins = 100, fill = "dark blue") +
  scale_x_log10(breaks = trans_breaks("log10", function(x) 10^x),
                labels = prettyNum0) + 
  theme(panel.grid.minor = element_blank()) +
  facet_wrap(~year) +
  labs(title = "Distribution of % applicants - TOP12 - method 1",
       #subtitle = "Applicant means s/o who puts the programcode anywhere in their 12 choices ",
       subtitle = "Method 1 selects the first year each applicant appears in the HSAPS data.",
       y = "# schools",
       x = "% applicants - x-scale is log10")

#################################################################################################

top1_apps_and_schools_method1 <-
  top1_applications_method1 %>%
  left_join(minimal_allschoolcodes, by=c("program_applied" = "pgcode")) %>%
  select(year = year.x,
         total_applicants,
         pgcode = program_applied,
         pgname,
         dbn,
         school_name,
         num_applications,
         app_rate,
         pc_applications)

top1_fooz_method1 <- 
  top1_apps_and_schools_method1 %>%
  filter(!is.na(dbn)) %>%
  group_by(year, dbn, school_name) %>%
  summarize(num_applications = sum(num_applications),
            total_applicants = mean(total_applicants),
            app_rate = num_applications/total_applicants)

# TOP 1
##################################################################################################
top1_applications_method1 %>%
  ggplot(aes(x=app_rate))+
  geom_histogram(bins = 100, fill = "blue") +
  scale_x_continuous(label = percent) + 
  theme(panel.grid.minor = element_blank()) +
  facet_wrap(~year) +
  labs(title = "Distribution of % applicants - TOP1 - method 1",
       #subtitle = "Applicant means s/o who puts the programcode in their FIRST choice",
       subtitle = "Method 1 selects the first year each applicant appears in the HSAPS data.",
       y = "# schools",
       x = "% applicants - x-scale is continous")

top1_applications_method1 %>%
  ggplot(aes(x=app_rate))+
  geom_histogram(bins = 100, fill = "blue") +
  scale_x_log10(breaks = trans_breaks("log10", function(x) 10^x),
                labels = prettyNum0) +  
  theme(panel.grid.minor = element_blank()) +
  facet_wrap(~year) +
  labs(title = "Distribution of % applicants - TOP1 - method 1",
       #subtitle = "Applicant means s/o who puts the programcode in their FIRST choice",
       subtitle = "Method 1 selects the first year each applicant appears in the HSAPS data.",
       y = "# schools",
       x = "% applicants - x-scale is log10")
```