---
title: '20170719'
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Loading in data.
```{r}
library(tidyverse)
library(ggplot2)
library(scales)
library(directlabels)
library(dplyr)

files <- Sys.glob('/data/nycdoe/Grad/*.csv')

read_csv_with_year <- function(filename) {
  year <- as.numeric(gsub('-.*$', '', basename(filename)))
  df <- read_csv(filename, col_types = cols(district_code=col_number(),
                                            school_resp_code=col_character(),
                                            dbn=col_character(),
                                            school_name=col_character()))
  df$year <- year
  df
}

grad <- map_df(files, read_csv_with_year)

grad %>%
  arrange(student_id_scram)%>%
  select(student_id_scram, membership_desc, class_of) %>%
  group_by(student_id_scram)%>%
  dplyr::summarise(count=n())%>%
  filter(count==1)
```

Is the 'graduated' bit consistent for each student?
```{r}
consistency_df <- 
  grad %>%
  group_by(student_id_scram) %>%
  summarise(is_consistent=length(unique(graduated))==1)

consistency_df %>%
  summarise(frac_consistent=mean(is_consistent))

(non_consistent_students <-
  grad %>%
  left_join(consistency_df, by=c("student_id_scram")) %>%
  filter(is_consistent==F) %>%
  select(student_id_scram, membership_desc, class_of, graduated) %>%
  arrange(student_id_scram) )

(consistent_students <-
  grad %>%
  left_join(consistency_df, by=c("student_id_scram")) %>%
  filter(is_consistent==T) %>%
  select(student_id_scram, membership_desc, class_of, graduated) %>%
  arrange(student_id_scram) )
```
89% of students have a consistent value for {r graduated}. For those students, we are assuming that the first (earliest) cohort listed is their cohort. For the other students, we are assuming that the first (earliest) cohort listed with a 1 for {r graduated} is their cohort.


```{r}
tidy_grad <-
  grad %>%
  group_by(student_id_scram) %>%
  summarise(graduated=max(graduated))
```

Loading in June Bios data (see 20170718.Rmd in Ilana's folder)
```{r}
library(tidyverse)
library(ggplot2)
library(scales)
library(directlabels)

files <- Sys.glob('/data/nycdoe/June Biog/*.csv')

read_csv_with_year <- function(filename) {
  year <- as.numeric(gsub('-.*$', '', basename(filename)))
  df <- read_csv(filename, col_types = cols(grade_level=col_character(), grade_code=col_character(), admit_code=col_character(), adcode=col_character(),fall_days_present = col_character(), spring_days_present = col_character()))
  df$year <- year
  df
}

bios <- map_df(files, read_csv_with_year) %>%
  filter(grade_level!="AD") %>%
  mutate(grade_level=ifelse(grade_level=="0K", 0, 
                            ifelse(grade_level=="PK", -1,
                                   ifelse(grade_level=="IN", -2, 
                                   as.numeric(grade_level)))))

df <- 
    bios %>%
    #filter(student_id_scram==420002000) %>%
    #arrange(student_id_scram) %>%
    select(student_id_scram, year, grade_level) %>%
    mutate(grade_level=as.numeric(grade_level)) %>%
    group_by(student_id_scram) %>%
    arrange(year) %>%
    mutate(did_continue=ifelse(!is.na(lead(grade_level)) & grade_level == lead(grade_level) - 1, 1, 0)) 

entries_per_student <-
    df %>%
    group_by(student_id_scram) %>%
    filter(ifelse(year==2013, grade_level>9, 
                  ifelse(year==2014, grade_level>10,
                         ifelse(year==2015, grade_level>11, grade_level>8))))
  
```

```{r}
new_df <-
  entries_per_student %>%
  left_join(tidy_grad, by=c('student_id_scram'))

new_df %>%
  ungroup() %>%
  summarise(p=sum(is.na(graduated))/n())
```

9.5% of the students in the June Bios data are missing from the Grad data.

```{r}
new_df_2 <-
  new_df %>%
  filter(!any(is.na(graduated)))
# %>%
#   filter(any(grade_level==9)) %>%
#   group_by(year) %>%
#   #distinct(student_id_scram) %>%
#   summarise(num_ninth=sum(grade_level==9, na.rm=T), 
#             continued_to_tenth=sum(grade_level==9 & did_continue==1, na.rm=T), 
#             continued_to_eleventh=sum(grade_level==10 & did_continue==1, na.rm=T), 
#             continued_to_twelfth=sum(grade_level==11 & did_continue==1, na.rm=T),
#             graduated=sum(grade_level==12 & graduated==1, na.rm=T),
#             perc_ninth=num_ninth/num_ninth,
#             perc_tenth=continued_to_tenth/num_ninth,
#             perc_eleventh=continued_to_eleventh/num_ninth,
#             perc_twelfth=continued_to_twelfth/num_ninth,
#             perc_graduated=graduated/num_ninth)

ninth <-
  new_df_2 %>%
  filter(any(grade_level==9))

tenth <-
  ninth %>%
  filter(any(grade_level==10))

eleventh <-
  tenth %>%
  filter(any(grade_level==11))

twelfth <-
  eleventh %>%
  filter(any(grade_level==12))

graduates <-
  twelfth %>%
  filter(max(graduated)==1)

a <-
  ninth %>%
  group_by(year) %>%
  distinct(student_id_scram) %>%
  dplyr::summarise(num_ninth=n())

b <-
  tenth %>%
  group_by(year) %>%
  distinct(student_id_scram) %>%
  dplyr::summarise(continued_to_tenth=n())

c <-
  eleventh %>%
  group_by(year) %>%
  distinct(student_id_scram) %>%
  dplyr::summarise(continued_to_eleventh=n())

d <-
  twelfth %>%
  group_by(year) %>%
  distinct(student_id_scram) %>%
  dplyr::summarise(continued_to_twelfth=n())

e <-
  graduates %>%
  group_by(year) %>%
  distinct(student_id_scram) %>%
  dplyr::summarise(graduated=n())

library(plyr)
library(dplyr)
final_df <-
  join_all(list(a, b, c, d, e), by='year', type='left')  
detach(package:plyr)

final_df_2 <-
  final_df %>%
  mutate(perc_ninth=num_ninth/num_ninth,
            perc_tenth=continued_to_tenth/num_ninth,
            perc_eleventh=continued_to_eleventh/num_ninth,
            perc_twelfth=continued_to_twelfth/num_ninth,
            perc_graduated=graduated/num_ninth) %>%
  select(year, 7:11) %>%
  gather("grade", "num_students", 2:6) %>%
  arrange(year)

final_df_2$grade[final_df_2$grade=="perc_ninth"] <- 9
final_df_2$grade[final_df_2$grade=="perc_tenth"] <- 10
final_df_2$grade[final_df_2$grade=="perc_eleventh"] <- 11
final_df_2$grade[final_df_2$grade=="perc_twelfth"] <- 12
final_df_2$grade[final_df_2$grade=="perc_graduated"] <- 13

final_df_2 <-
  final_df_2 %>%
  mutate(grade=as.numeric(grade))

final_df_2

final_df_2 %>%
  filter(year<=2012) %>%
  ggplot(aes(x=grade, y=num_students, color=as.factor(year))) +
  geom_point() +
  geom_line() +
  scale_y_continuous(labels = percent) +
  scale_x_continuous(breaks=seq(9,13,1), labels=c('9'='9', '10'='10', '11'='11', '12'='12', '13'='Graduated'), expand=c(0, 0.75)) +
  geom_dl(aes(label = as.factor(year)), method = list(dl.trans(x = x + .3), "last.points", cex = 0.8)) +
  labs(x="Grade Level", y="Percent of Incoming Freshmen", title="Student Drop-Outs Throughout High School", color="Class of")
```

Facet by ethnicity.

```{r}
students_with_eth <-
  bios %>%
  select(student_id_scram, ethnicity) %>%
  distinct(student_id_scram, ethnicity) %>%
  filter(!is.na(student_id_scram))

new_df_3 <-
  new_df_2 %>%
  left_join(students_with_eth, by='student_id_scram')

ninth <-
  new_df_3 %>%
  filter(any(grade_level==9))

tenth <-
  ninth %>%
  filter(any(grade_level==10))

eleventh <-
  tenth %>%
  filter(any(grade_level==11))

twelfth <-
  eleventh %>%
  filter(any(grade_level==12))

graduates <-
  twelfth %>%
  filter(max(graduated)==1)

a <-
  ninth %>%
  group_by(ethnicity, year) %>%
  distinct(student_id_scram) %>%
  dplyr::summarise(num_ninth=n())

b <-
  tenth %>%
  group_by(ethnicity, year) %>%
  distinct(student_id_scram) %>%
  dplyr::summarise(continued_to_tenth=n())

c <-
  eleventh %>%
  group_by(ethnicity, year) %>%
  distinct(student_id_scram) %>%
  dplyr::summarise(continued_to_eleventh=n())

d <-
  twelfth %>%
  group_by(ethnicity, year) %>%
  distinct(student_id_scram) %>%
  dplyr::summarise(continued_to_twelfth=n())

e <-
  graduates %>%
  group_by(ethnicity, year) %>%
  distinct(student_id_scram) %>%
  dplyr::summarise(graduated=n())

# library(plyr)
# library(dplyr)
# final_df <-
#   join_all(list(a, b, c, d, e), by=c('ethnicity','year'), type='left')  
# detach(package:plyr)

final_df <-
  a %>%
  left_join(b, by=c('ethnicity', 'year')) %>%
  left_join(c, by=c('ethnicity', 'year')) %>%
  left_join(d, by=c('ethnicity', 'year')) %>%
  left_join(e, by=c('ethnicity', 'year'))

final_df_2 <-
  final_df %>%
  mutate(perc_ninth=num_ninth/num_ninth,
            perc_tenth=continued_to_tenth/num_ninth,
            perc_eleventh=continued_to_eleventh/num_ninth,
            perc_twelfth=continued_to_twelfth/num_ninth,
            perc_graduated=graduated/num_ninth) %>%
  select(ethnicity, year, 8:12) %>%
  gather("grade", "num_students", 3:7) %>%
  arrange(year)

final_df_2$grade[final_df_2$grade=="perc_ninth"] <- 9
final_df_2$grade[final_df_2$grade=="perc_tenth"] <- 10
final_df_2$grade[final_df_2$grade=="perc_eleventh"] <- 11
final_df_2$grade[final_df_2$grade=="perc_twelfth"] <- 12
final_df_2$grade[final_df_2$grade=="perc_graduated"] <- 13

final_df_2 <-
  final_df_2 %>%
  mutate(grade=as.numeric(grade))

final_df_2

final_df_2 %>%
  filter(year<=2012) %>%
  ggplot(aes(x=grade, y=num_students, color=as.factor(year))) +
  geom_point() +
  geom_line() +
  scale_y_continuous(labels = percent) +
  scale_x_continuous(breaks=seq(9,13,1), labels=c('9'='9', '10'='10', '11'='11', '12'='12', '13'='Graduated'), expand=c(0, 0.75)) +
  geom_dl(aes(label = as.factor(year)), method = list(dl.trans(x = x + .3), "last.points", cex = 0.8)) +
  labs(x="Grade Level", y="Percent of Incoming Freshmen", title="Student Drop-Outs Throughout High School", color="Class of") +
  facet_wrap(~ethnicity)
```

Facet by borough.

```{r}
students_with_boro <-
  bios %>%
  select(student_id_scram, dbn) %>%
  mutate(boro=substring(dbn, 3, 3)) %>%
  distinct(student_id_scram, boro) %>%
  filter(!is.na(student_id_scram))

new_df_3 <-
  new_df_2 %>%
  left_join(students_with_boro, by='student_id_scram')

ninth <-
  new_df_3 %>%
  filter(any(grade_level==9))

tenth <-
  ninth %>%
  filter(any(grade_level==10))

eleventh <-
  tenth %>%
  filter(any(grade_level==11))

twelfth <-
  eleventh %>%
  filter(any(grade_level==12))

graduates <-
  twelfth %>%
  filter(max(graduated)==1)

a <-
  ninth %>%
  group_by(boro, year) %>%
  distinct(student_id_scram) %>%
  dplyr::summarise(num_ninth=n())

b <-
  tenth %>%
  group_by(boro, year) %>%
  distinct(student_id_scram) %>%
  dplyr::summarise(continued_to_tenth=n())

c <-
  eleventh %>%
  group_by(boro, year) %>%
  distinct(student_id_scram) %>%
  dplyr::summarise(continued_to_eleventh=n())

d <-
  twelfth %>%
  group_by(boro, year) %>%
  distinct(student_id_scram) %>%
  dplyr::summarise(continued_to_twelfth=n())

e <-
  graduates %>%
  group_by(boro, year) %>%
  distinct(student_id_scram) %>%
  dplyr::summarise(graduated=n())

# library(plyr)
# library(dplyr)
# final_df <-
#   join_all(list(a, b, c, d, e), by=c('ethnicity','year'), type='left')  
# detach(package:plyr)

final_df <-
  a %>%
  left_join(b, by=c('boro', 'year')) %>%
  left_join(c, by=c('boro', 'year')) %>%
  left_join(d, by=c('boro', 'year')) %>%
  left_join(e, by=c('boro', 'year'))

final_df_2 <-
  final_df %>%
  mutate(perc_ninth=num_ninth/num_ninth,
            perc_tenth=continued_to_tenth/num_ninth,
            perc_eleventh=continued_to_eleventh/num_ninth,
            perc_twelfth=continued_to_twelfth/num_ninth,
            perc_graduated=graduated/num_ninth) %>%
  select(boro, year, 8:12) %>%
  gather("grade", "num_students", 3:7) %>%
  arrange(year)

final_df_2$grade[final_df_2$grade=="perc_ninth"] <- 9
final_df_2$grade[final_df_2$grade=="perc_tenth"] <- 10
final_df_2$grade[final_df_2$grade=="perc_eleventh"] <- 11
final_df_2$grade[final_df_2$grade=="perc_twelfth"] <- 12
final_df_2$grade[final_df_2$grade=="perc_graduated"] <- 13

final_df_2 <-
  final_df_2 %>%
  mutate(grade=as.numeric(grade))

final_df_2

final_df_2 %>%
  filter(year<=2012) %>%
  ggplot(aes(x=grade, y=num_students, color=as.factor(year))) +
  geom_point() +
  geom_line() +
  scale_y_continuous(labels = percent) +
  scale_x_continuous(breaks=seq(9,13,1), labels=c('9'='9', '10'='10', '11'='11', '12'='12', '13'='Graduated'), expand=c(0, 0.75)) +
  geom_dl(aes(label = as.factor(year)), method = list(dl.trans(x = x + .3), "last.points", cex = 0.8)) +
  labs(x="Grade Level", y="Percent of Incoming Freshmen", title="Student Drop-Outs Throughout High School", color="Class of") +
  facet_wrap(~boro)
```

Facet by us-born.

```{r}
students_with_pob <-
  bios %>%
  select(student_id_scram, pob_code) %>%
  mutate(us_born=!is.na(as.numeric(pob_code))) %>%
  distinct(student_id_scram, us_born) %>%
  filter(!is.na(student_id_scram))

new_df_3 <-
  new_df_2 %>%
  left_join(students_with_pob, by='student_id_scram')

ninth <-
  new_df_3 %>%
  filter(any(grade_level==9))

tenth <-
  ninth %>%
  filter(any(grade_level==10))

eleventh <-
  tenth %>%
  filter(any(grade_level==11))

twelfth <-
  eleventh %>%
  filter(any(grade_level==12))

graduates <-
  twelfth %>%
  filter(max(graduated)==1)

a <-
  ninth %>%
  group_by(us_born, year) %>%
  distinct(student_id_scram) %>%
  dplyr::summarise(num_ninth=n())

b <-
  tenth %>%
  group_by(us_born, year) %>%
  distinct(student_id_scram) %>%
  dplyr::summarise(continued_to_tenth=n())

c <-
  eleventh %>%
  group_by(us_born, year) %>%
  distinct(student_id_scram) %>%
  dplyr::summarise(continued_to_eleventh=n())

d <-
  twelfth %>%
  group_by(us_born, year) %>%
  distinct(student_id_scram) %>%
  dplyr::summarise(continued_to_twelfth=n())

e <-
  graduates %>%
  group_by(us_born, year) %>%
  distinct(student_id_scram) %>%
  dplyr::summarise(graduated=n())

# library(plyr)
# library(dplyr)
# final_df <-
#   join_all(list(a, b, c, d, e), by=c('ethnicity','year'), type='left')  
# detach(package:plyr)

final_df <-
  a %>%
  left_join(b, by=c('us_born', 'year')) %>%
  left_join(c, by=c('us_born', 'year')) %>%
  left_join(d, by=c('us_born', 'year')) %>%
  left_join(e, by=c('us_born', 'year'))

final_df_2 <-
  final_df %>%
  mutate(perc_ninth=num_ninth/num_ninth,
            perc_tenth=continued_to_tenth/num_ninth,
            perc_eleventh=continued_to_eleventh/num_ninth,
            perc_twelfth=continued_to_twelfth/num_ninth,
            perc_graduated=graduated/num_ninth) %>%
  select(us_born, year, 8:12) %>%
  gather("grade", "num_students", 3:7) %>%
  arrange(year)

final_df_2$grade[final_df_2$grade=="perc_ninth"] <- 9
final_df_2$grade[final_df_2$grade=="perc_tenth"] <- 10
final_df_2$grade[final_df_2$grade=="perc_eleventh"] <- 11
final_df_2$grade[final_df_2$grade=="perc_twelfth"] <- 12
final_df_2$grade[final_df_2$grade=="perc_graduated"] <- 13

final_df_2 <-
  final_df_2 %>%
  mutate(grade=as.numeric(grade))

final_df_2

final_df_2 %>%
  filter(year<=2012) %>%
  ggplot(aes(x=grade, y=num_students, color=as.factor(year))) +
  geom_point() +
  geom_line() +
  scale_y_continuous(labels = percent) +
  scale_x_continuous(breaks=seq(9,13,1), labels=c('9'='9', '10'='10', '11'='11', '12'='12', '13'='Graduated'), expand=c(0, 0.75)) +
  geom_dl(aes(label = as.factor(year)), method = list(dl.trans(x = x + .3), "last.points", cex = 0.8)) +
  labs(x="Grade Level", y="Percent of Incoming Freshmen", title="Student Drop-Outs Throughout High School", color="Class of") +
  facet_wrap(~us_born)
```

Facet by borough and us_born.

```{r}
students_with_boro_and_us <-
  bios %>%
  select(student_id_scram, dbn, pob_code) %>%
  mutate(boro=substring(dbn, 3, 3), us_born=!is.na(as.numeric(pob_code))) %>%
  distinct(student_id_scram, boro, us_born) %>%
  filter(!is.na(student_id_scram))

new_df_3 <-
  new_df_2 %>%
  left_join(students_with_boro_and_us, by='student_id_scram')

ninth <-
  new_df_3 %>%
  filter(any(grade_level==9))

tenth <-
  ninth %>%
  filter(any(grade_level==10))

eleventh <-
  tenth %>%
  filter(any(grade_level==11))

twelfth <-
  eleventh %>%
  filter(any(grade_level==12))

graduates <-
  twelfth %>%
  filter(max(graduated)==1)

a <-
  ninth %>%
  group_by(boro, us_born, year) %>%
  distinct(student_id_scram) %>%
  dplyr::summarise(num_ninth=n())

b <-
  tenth %>%
  group_by(boro, us_born, year) %>%
  distinct(student_id_scram) %>%
  dplyr::summarise(continued_to_tenth=n())

c <-
  eleventh %>%
  group_by(boro, us_born, year) %>%
  distinct(student_id_scram) %>%
  dplyr::summarise(continued_to_eleventh=n())

d <-
  twelfth %>%
  group_by(boro, us_born, year) %>%
  distinct(student_id_scram) %>%
  dplyr::summarise(continued_to_twelfth=n())

e <-
  graduates %>%
  group_by(boro, us_born, year) %>%
  distinct(student_id_scram) %>%
  dplyr::summarise(graduated=n())

# library(plyr)
# library(dplyr)
# final_df <-
#   join_all(list(a, b, c, d, e), by=c('ethnicity','year'), type='left')  
# detach(package:plyr)

final_df <-
  a %>%
  left_join(b, by=c('boro', 'us_born', 'year')) %>%
  left_join(c, by=c('boro', 'us_born', 'year')) %>%
  left_join(d, by=c('boro', 'us_born', 'year')) %>%
  left_join(e, by=c('boro', 'us_born', 'year'))

final_df_2 <-
  final_df %>%
  mutate(perc_ninth=num_ninth/num_ninth,
            perc_tenth=continued_to_tenth/num_ninth,
            perc_eleventh=continued_to_eleventh/num_ninth,
            perc_twelfth=continued_to_twelfth/num_ninth,
            perc_graduated=graduated/num_ninth) %>%
  select(boro, us_born, year, 9:13) %>%
  gather("grade", "num_students", 4:8) %>%
  arrange(year)

final_df_2$grade[final_df_2$grade=="perc_ninth"] <- 9
final_df_2$grade[final_df_2$grade=="perc_tenth"] <- 10
final_df_2$grade[final_df_2$grade=="perc_eleventh"] <- 11
final_df_2$grade[final_df_2$grade=="perc_twelfth"] <- 12
final_df_2$grade[final_df_2$grade=="perc_graduated"] <- 13

final_df_2 <-
  final_df_2 %>%
  mutate(grade=as.numeric(grade))

final_df_2

final_df_2 %>%
  filter(year<=2012) %>%
  ggplot(aes(x=grade, y=num_students, color=as.factor(year))) +
  geom_point() +
  geom_line() +
  scale_y_continuous(labels = percent) +
  scale_x_continuous(breaks=seq(9,13,1), labels=c('9'='9', '10'='10', '11'='11', '12'='12', '13'='Graduated'), expand=c(0, 0.75)) +
  geom_dl(aes(label = as.factor(year)), method = list(dl.trans(x = x + .3), "last.points", cex = 0.8)) +
  labs(x="Grade Level", y="Percent of Incoming Freshmen", title="Student Drop-Outs Throughout High School", color="Class of") +
  facet_grid(us_born~boro)
```

Facet by gender.

```{r}
students_with_sex <-
  bios %>%
  select(student_id_scram, sex) %>%
  distinct(student_id_scram, sex) %>%
  filter(!is.na(student_id_scram))

new_df_3 <-
  new_df_2 %>%
  left_join(students_with_sex, by='student_id_scram')

ninth <-
  new_df_3 %>%
  filter(any(grade_level==9))

tenth <-
  ninth %>%
  filter(any(grade_level==10))

eleventh <-
  tenth %>%
  filter(any(grade_level==11))

twelfth <-
  eleventh %>%
  filter(any(grade_level==12))

graduates <-
  twelfth %>%
  filter(max(graduated)==1)

a <-
  ninth %>%
  group_by(sex, year) %>%
  distinct(student_id_scram) %>%
  dplyr::summarise(num_ninth=n())

b <-
  tenth %>%
  group_by(sex, year) %>%
  distinct(student_id_scram) %>%
  dplyr::summarise(continued_to_tenth=n())

c <-
  eleventh %>%
  group_by(sex, year) %>%
  distinct(student_id_scram) %>%
  dplyr::summarise(continued_to_eleventh=n())

d <-
  twelfth %>%
  group_by(sex, year) %>%
  distinct(student_id_scram) %>%
  dplyr::summarise(continued_to_twelfth=n())

e <-
  graduates %>%
  group_by(sex, year) %>%
  distinct(student_id_scram) %>%
  dplyr::summarise(graduated=n())

# library(plyr)
# library(dplyr)
# final_df <-
#   join_all(list(a, b, c, d, e), by=c('ethnicity','year'), type='left')  
# detach(package:plyr)

final_df <-
  a %>%
  left_join(b, by=c('sex', 'year')) %>%
  left_join(c, by=c('sex', 'year')) %>%
  left_join(d, by=c('sex', 'year')) %>%
  left_join(e, by=c('sex', 'year'))

final_df_2 <-
  final_df %>%
  mutate(perc_ninth=num_ninth/num_ninth,
            perc_tenth=continued_to_tenth/num_ninth,
            perc_eleventh=continued_to_eleventh/num_ninth,
            perc_twelfth=continued_to_twelfth/num_ninth,
            perc_graduated=graduated/num_ninth) %>%
  select(sex, year, 8:12) %>%
  gather("grade", "num_students", 3:7) %>%
  arrange(year)

final_df_2$grade[final_df_2$grade=="perc_ninth"] <- 9
final_df_2$grade[final_df_2$grade=="perc_tenth"] <- 10
final_df_2$grade[final_df_2$grade=="perc_eleventh"] <- 11
final_df_2$grade[final_df_2$grade=="perc_twelfth"] <- 12
final_df_2$grade[final_df_2$grade=="perc_graduated"] <- 13

final_df_2 <-
  final_df_2 %>%
  mutate(grade=as.numeric(grade))

final_df_2

final_df_2 %>%
  filter(year<=2012) %>%
  ggplot(aes(x=grade, y=num_students, color=as.factor(year))) +
  geom_point() +
  geom_line() +
  scale_y_continuous(labels = percent) +
  scale_x_continuous(breaks=seq(9,13,1), labels=c('9'='9', '10'='10', '11'='11', '12'='12', '13'='Graduated'), expand=c(0, 0.75)) +
  geom_dl(aes(label = as.factor(year)), method = list(dl.trans(x = x + .3), "last.points", cex = 0.8)) +
  labs(x="Grade Level", y="Percent of Incoming Freshmen", title="Student Drop-Outs Throughout High School", color="Class of") +
  facet_wrap(~sex)
```

Facet by poverty.

```{r}
students_with_poverty <-
  bios %>%
  select(student_id_scram, poverty) %>%
  distinct(student_id_scram, poverty) %>%
  filter(!is.na(student_id_scram))

new_df_3 <-
  new_df_2 %>%
  left_join(students_with_poverty, by='student_id_scram')

ninth <-
  new_df_3 %>%
  filter(any(grade_level==9))

tenth <-
  ninth %>%
  filter(any(grade_level==10))

eleventh <-
  tenth %>%
  filter(any(grade_level==11))

twelfth <-
  eleventh %>%
  filter(any(grade_level==12))

graduates <-
  twelfth %>%
  filter(max(graduated)==1)

a <-
  ninth %>%
  group_by(poverty, year) %>%
  distinct(student_id_scram) %>%
  dplyr::summarise(num_ninth=n())

b <-
  tenth %>%
  group_by(poverty, year) %>%
  distinct(student_id_scram) %>%
  dplyr::summarise(continued_to_tenth=n())

c <-
  eleventh %>%
  group_by(poverty, year) %>%
  distinct(student_id_scram) %>%
  dplyr::summarise(continued_to_eleventh=n())

d <-
  twelfth %>%
  group_by(poverty, year) %>%
  distinct(student_id_scram) %>%
  dplyr::summarise(continued_to_twelfth=n())

e <-
  graduates %>%
  group_by(poverty, year) %>%
  distinct(student_id_scram) %>%
  dplyr::summarise(graduated=n())

# library(plyr)
# library(dplyr)
# final_df <-
#   join_all(list(a, b, c, d, e), by=c('ethnicity','year'), type='left')  
# detach(package:plyr)

final_df <-
  a %>%
  left_join(b, by=c('poverty', 'year')) %>%
  left_join(c, by=c('poverty', 'year')) %>%
  left_join(d, by=c('poverty', 'year')) %>%
  left_join(e, by=c('poverty', 'year'))

final_df_2 <-
  final_df %>%
  mutate(perc_ninth=num_ninth/num_ninth,
            perc_tenth=continued_to_tenth/num_ninth,
            perc_eleventh=continued_to_eleventh/num_ninth,
            perc_twelfth=continued_to_twelfth/num_ninth,
            perc_graduated=graduated/num_ninth) %>%
  select(poverty, year, 8:12) %>%
  gather("grade", "num_students", 3:7) %>%
  arrange(year)

final_df_2$grade[final_df_2$grade=="perc_ninth"] <- 9
final_df_2$grade[final_df_2$grade=="perc_tenth"] <- 10
final_df_2$grade[final_df_2$grade=="perc_eleventh"] <- 11
final_df_2$grade[final_df_2$grade=="perc_twelfth"] <- 12
final_df_2$grade[final_df_2$grade=="perc_graduated"] <- 13

final_df_2 <-
  final_df_2 %>%
  mutate(grade=as.numeric(grade))

final_df_2

final_df_2 %>%
  filter(year<=2012) %>%
  ggplot(aes(x=grade, y=num_students, color=as.factor(year))) +
  geom_point() +
  geom_line() +
  scale_y_continuous(labels = percent) +
  scale_x_continuous(breaks=seq(9,13,1), labels=c('9'='9', '10'='10', '11'='11', '12'='12', '13'='Graduated'), expand=c(0, 0.75)) +
  geom_dl(aes(label = as.factor(year)), method = list(dl.trans(x = x + .3), "last.points", cex = 0.8)) +
  labs(x="Grade Level", y="Percent of Incoming Freshmen", title="Student Drop-Outs Throughout High School", color="Class of") +
  facet_wrap(~poverty)
```

Facet by disability.

```{r}
students_with_disability <-
  bios %>%
  select(student_id_scram, disability) %>%
  mutate(disability=(!(is.na(disability)|disability=='ND'))) %>%
  distinct(student_id_scram, disability) %>%
  filter(!is.na(student_id_scram))

new_df_3 <-
  new_df_2 %>%
  left_join(students_with_disability, by='student_id_scram')

ninth <-
  new_df_3 %>%
  filter(any(grade_level==9))

tenth <-
  ninth %>%
  filter(any(grade_level==10))

eleventh <-
  tenth %>%
  filter(any(grade_level==11))

twelfth <-
  eleventh %>%
  filter(any(grade_level==12))

graduates <-
  twelfth %>%
  filter(max(graduated)==1)

a <-
  ninth %>%
  group_by(disability, year) %>%
  distinct(student_id_scram) %>%
  dplyr::summarise(num_ninth=n())

b <-
  tenth %>%
  group_by(disability, year) %>%
  distinct(student_id_scram) %>%
  dplyr::summarise(continued_to_tenth=n())

c <-
  eleventh %>%
  group_by(disability, year) %>%
  distinct(student_id_scram) %>%
  dplyr::summarise(continued_to_eleventh=n())

d <-
  twelfth %>%
  group_by(disability, year) %>%
  distinct(student_id_scram) %>%
  dplyr::summarise(continued_to_twelfth=n())

e <-
  graduates %>%
  group_by(disability, year) %>%
  distinct(student_id_scram) %>%
  dplyr::summarise(graduated=n())

# library(plyr)
# library(dplyr)
# final_df <-
#   join_all(list(a, b, c, d, e), by=c('ethnicity','year'), type='left')  
# detach(package:plyr)

final_df <-
  a %>%
  left_join(b, by=c('disability', 'year')) %>%
  left_join(c, by=c('disability', 'year')) %>%
  left_join(d, by=c('disability', 'year')) %>%
  left_join(e, by=c('disability', 'year'))

final_df_2 <-
  final_df %>%
  mutate(perc_ninth=num_ninth/num_ninth,
            perc_tenth=continued_to_tenth/num_ninth,
            perc_eleventh=continued_to_eleventh/num_ninth,
            perc_twelfth=continued_to_twelfth/num_ninth,
            perc_graduated=graduated/num_ninth) %>%
  select(disability, year, 8:12) %>%
  gather("grade", "num_students", 3:7) %>%
  arrange(year)

final_df_2$grade[final_df_2$grade=="perc_ninth"] <- 9
final_df_2$grade[final_df_2$grade=="perc_tenth"] <- 10
final_df_2$grade[final_df_2$grade=="perc_eleventh"] <- 11
final_df_2$grade[final_df_2$grade=="perc_twelfth"] <- 12
final_df_2$grade[final_df_2$grade=="perc_graduated"] <- 13

final_df_2 <-
  final_df_2 %>%
  mutate(grade=as.numeric(grade))

final_df_2

final_df_2 %>%
  filter(year<=2012) %>%
  ggplot(aes(x=grade, y=num_students, color=as.factor(year))) +
  geom_point() +
  geom_line() +
  scale_y_continuous(labels = percent) +
  scale_x_continuous(breaks=seq(9,13,1), labels=c('9'='9', '10'='10', '11'='11', '12'='12', '13'='Graduated'), expand=c(0, 0.75)) +
  geom_dl(aes(label = as.factor(year)), method = list(dl.trans(x = x + .3), "last.points", cex = 0.8)) +
  labs(x="Grade Level", y="Percent of Incoming Freshmen", title="Student Drop-Outs Throughout High School", color="Class of") +
  facet_wrap(~disability)
```
