---
title: "School Demand/Popularity"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(tidyverse)
library(ggplot2)
library(scales)

files <- Sys.glob('/data/nycdoe/HSAPS/*.csv')

read_students_with_choices <- function(filename) {
  year <- as.numeric(gsub('-.*$', '', basename(filename)))
  df <- read_csv(filename, col_types = cols(student_id_scram=col_character()))
  df$year <- year
  df <- df %>% 
    select(student_id_scram, contains("programcode"), year)
}

tidy_df <-
  map_df(files, read_students_with_choices) %>%
  filter(!is.na(student_id_scram)) %>%
  gather("round_choice", "school", 2:37) %>%
  mutate(round = substring(round_choice, 2,2), 
         choice = substring(round_choice, 14)) %>%
  select(-round_choice) %>%
  mutate(student_id_scram = as.factor(student_id_scram),
         school = as.factor(school))
```

How many people ranked each school?
```{r}
num_apps_per_school <-
  tidy_df %>%
  group_by(school) %>%
  summarise(num_apps=n())

head(num_apps_per_school)
```

```{r}
num_apps_per_school_w_round <-
  tidy_df %>%
  group_by(school, round) %>%
  summarise(num_apps=n())
head(num_apps_per_school_w_round)
```
# Plots

## Top 20 most popular schools - number of applications

```{r}
num_apps_per_school_w_round %>%
  filter(!is.na(school)) %>%
  arrange(desc(num_apps)) %>%
  group_by(round) %>%
  head(20) %>%
  arrange(num_apps) %>%
  ggplot(aes(x=num_apps, y=factor(school, levels=unique(as.character(school)) ))) +
  geom_point()
```

## Distribution of num apps over all schools

```{r}
num_apps_per_school_w_round %>%
  filter(!is.na(school)) %>%
  group_by(num_apps, round) %>%
  summarise(count=n()) %>%
  ggplot(aes(x=num_apps, y=count)) +
  geom_line() +
  scale_x_log10(label=comma) +
  facet_wrap(~round)
```

```{r}
num_apps_per_school_w_round %>%
  filter(!is.na(school)) %>%
  group_by(num_apps, round) %>%
  summarise(count=n()) %>%
  ggplot(aes(x=num_apps)) +
  geom_density(color = "purple", fill="purple", alpha=0.2) +
  facet_wrap(~round, scale="free_y")
```

## Where are schools typically listed in the rankings?

Showing the top 100 schools with the most applications

```{r}
tidy_df %>%
  filter(!is.na(school)) %>%
  group_by(school, round) %>%
  mutate(choice=as.numeric(choice)) %>%
  summarise(avg_rank=mean(choice, na.rm=T), se = sd(choice, na.rm=T)/sqrt(n()), num_apps=n()) %>%
  #arrange(se, avg_rank) %>%
  arrange(desc(num_apps)) %>%
  head(100) %>%
  ggplot(aes(y=avg_rank, x=factor(school, levels=unique(as.character(school))), color=num_apps)) +
  geom_point() +
  scale_colour_gradient(low = "red", high = "purple") +
  geom_pointrange(aes(ymin=avg_rank-se, ymax=avg_rank+se)) +
  coord_flip()
```

Showing all schools
```{r}
tidy_df %>%
  filter(!is.na(school)) %>%
  group_by(school, round) %>%
  mutate(choice=as.numeric(choice)) %>%
  summarise(avg_rank=mean(choice, na.rm=T), se = sd(choice, na.rm=T)/sqrt(n()), num_apps=n()) %>%
  arrange(avg_rank) %>%
  ggplot(aes(y=avg_rank, x=factor(school, levels=unique(as.character(school))), color=num_apps)) +
  geom_point(alpha=0.5) +
  scale_colour_gradient(low = "red", high = "purple") +
  geom_pointrange(aes(ymin=avg_rank-se, ymax=avg_rank+se)) +
  coord_flip() +
  facet_wrap(~round)
```

Switched out school names for popularity (schools with highest avg rank -> school_pop=1)
```{r}
tidy_df %>%
  filter(!is.na(school)) %>%
  group_by(school, round) %>%
  mutate(choice=as.numeric(choice)) %>%
  summarise(avg_rank=mean(choice, na.rm=T), se = sd(choice, na.rm=T)/sqrt(n()), num_apps=n()) %>%
  arrange(avg_rank) %>%
  ungroup() %>%
  #changed school_popularity based on num_apps instead of average rank
  mutate(school_pop=rank(desc(num_apps))) %>%
  ggplot(aes(y=avg_rank, x=school_pop, color=num_apps)) +
  geom_point() +
  scale_color_gradientn(colors = rainbow(4)) +
  geom_pointrange(aes(ymin=avg_rank-se, ymax=avg_rank+se)) +
  facet_wrap(~round, ncol=1) + 
  scale_x_log10()
```

Same thing with geom_smooth()
```{r}
tidy_df %>%
  filter(!is.na(school)) %>%
  group_by(school, round) %>%
  mutate(choice=as.numeric(choice)) %>%
  summarise(avg_rank=mean(choice, na.rm=T), se = sd(choice, na.rm=T)/sqrt(n()), num_apps=n()) %>%
  arrange(avg_rank) %>%
  ungroup() %>%
  mutate(school_pop=row_number()) %>%
  ggplot(aes(y=avg_rank, x=school_pop)) +
  geom_smooth() +
  coord_flip() +
  facet_wrap(~round)
```

Distribution of mean ranks by school
```{r}
tidy_df %>%
  filter(!is.na(school)) %>%
  group_by(school, round) %>%
  mutate(choice=as.numeric(choice)) %>%
  summarise(avg_rank=mean(choice, na.rm=T)) %>%
  ggplot(aes(x=avg_rank)) +
  geom_density(color="blue", fill="blue", alpha=0.2) +
  facet_wrap(~round)
```

Distribution of mean ranks vs number of applications
(Each school is one point)
```{r}
tidy_df %>%
  filter(!is.na(school)) %>%
  group_by(school, round) %>%
  mutate(choice=as.numeric(choice)) %>%
  summarise(avg_rank=mean(choice, na.rm=T), count=n()) %>%
  ggplot(aes(x=avg_rank, y=count)) +
  geom_point() +
  facet_wrap(~round)
```

```{r}
tidy_df %>%
  filter(!is.na(school)) %>%
  group_by(school, round) %>%
  mutate(choice=as.numeric(choice)) %>%
  summarise(avg_rank=mean(choice, na.rm=T), count=n()) %>%
  ggplot(aes(x=avg_rank, y=count)) +
  geom_area(fill="orange", alpha=0.7) +
  facet_wrap(~round)
```








