---
title: '0719_typical_profile'
author: "Keri and David"
date: "7/18/2017"
output: html_document
---
Setting up. Function to read in multiple files, selecting:
Year, Student ID, DBN, Grade_Level, and Date
```{r setup}
knitr::opts_chunk$set(echo = TRUE, root.dir = '/data/nycdoe/')
options(tibble.width=Inf)

library(ggplot2)
library(readr)
library(tidyverse)
library(reshape2)
files <- Sys.glob('/data/nycdoe/June Biog/*.csv')

read_csv_with_year <- function(filename) {
  year <- as.numeric(gsub('-.*$', '', basename(filename)))
  df <- read_csv(filename, col_types = cols(student_id_scram = col_character(), grade_level = col_character()))
  df$year <- year
  df <- select(df, year, student_id_scram, birth_mm_yyyy, sex, ethnicity, dbn, grade_level, contains("date"))
}

```
Calling in Data
```{r using bios df}
setwd('/data/nycdoe/')
files <- Sys.glob('/data/nycdoe/June Biog/*.csv')
bios_backup <- map_df(files, read_csv_with_year)
bios <- bios_backup
```
Configuring data
```{r pressure, echo=TRUE}
bios <- bios %>%
  mutate(date = ifelse(!is.na(admit_date), admit_date, date)) %>%
  select(-contains("_date")) %>%
  mutate(date = floor(date/10000))
#Change birth_mm_yyyy to dob
colnames(bios)[3] <- "dob"
```
Convertingg grade level to numeric
Dropping anyone with an age under 3 as well as anyone with an age above 21 in grade under 9. we assume those are mistakes
```{r, echo=TRUE}
bios <- bios %>%
  filter(grade_level!='AD') %>%
  mutate(grade_level = ifelse(grade_level=='IN', '-2', grade_level)) %>%
  mutate(grade_level = ifelse(grade_level=='0K', '0', grade_level)) %>%
  mutate(grade_level = ifelse(grade_level=='PK', '-1', grade_level)) %>%
  mutate(grade_level = as.numeric(grade_level)) %>%
  mutate(dob = substr(dob, 4, 7)) %>%
  mutate(age = year - as.numeric(dob)) %>%
  filter(!(age<3))%>%
  filter(!(age>21 & grade_level<9))%>%
  select(year, student_id_scram, age, sex, ethnicity, dbn, grade_level, date)
```
Compare retention rate(?) of 5 of the best versus 5 of the worse schools
```{r}
best_vs_worst <- bios %>%
  filter(dbn == '10X696' | dbn == '02M542' | dbn == '05M692' | dbn == '31R605' | dbn == '02M545' | dbn == '02M460' | dbn == '07X547' | dbn == '14K22' | dbn == '24Q455' | dbn == '08x560') %>%
  mutate(label = ifelse(dbn == '10X696' | dbn == '02M542' | dbn == '05M692' | dbn == '31R605' | dbn == '02M545', "'Good' School", "'Bad' School")) %>%
  group_by(student_id_scram, label) %>% summarize(count = n()) %>%
  ggplot(aes(x = count)) +
  geom_bar(fill="blue") +
  facet_wrap(~label) +
  xlab('Number of years in the system') +
  ylab('Number of students')
plot(best_vs_worst)

```
Age vs. Grade between two schools
```{r}
two_schools <- 
  bios %>% 
  filter(dbn == '10X696' | dbn == '02M460')  %>%
  mutate(label = ifelse(dbn == '10X696', "'Good' School", "'Bad' School")) %>%
  select(year,dbn,label,student_id_scram,grade_level,age)
two_schools_plot <-
  two_schools %>%
  ggplot(aes(x=age, y=grade_level, group = student_id_scram, colour = student_id_scram)) +
  geom_line(size=1, show.legend=F) +
  geom_point(size=1, shape=21, fill="white", show.legend=F ) + 
  facet_wrap(~label)

#Special case: Student 302122382
#s82 <- bios %>% filter(student_id_scram == '02122382')
```
Age vs. Grade for ten schools. 
Getting a small sample of a 100 students
```{r}
ten_schools <- 
  bios %>%
  filter(dbn == '10X696' | dbn == '02M542' | dbn == '05M692' | dbn == '31R605' | dbn == '02M545' | dbn == '02M460' | dbn == '07X547' | dbn == '14K22' | dbn == '24Q455' | dbn == '08x560') %>%
  mutate(label = ifelse(dbn == '10X696' | dbn == '02M542' | dbn == '05M692' | dbn == '31R605' | dbn == '02M545',"'Good' School", "'Bad' School"))

set.seed(11)
tiny_sample_name <- ten_schools %>% select(student_id_scram,label) %>% unique() %>% group_by(label) %>% sample_n(100) %>% ungroup() %>% select(student_id_scram)
tiny_sample <- left_join(tiny_sample_name,ten_schools, by=c("student_id_scram"))
```
Plotting 100 student sample
```{r}
tiny_sample %>%
  ggplot(aes(x=age, y=grade_level, group = student_id_scram, colour = student_id_scram)) +
  geom_line(size=1, show.legend=F) +
  geom_point(size=1, shape=21, fill="white", show.legend=F ) +
  facet_wrap(~label) +scale_x_continuous(breaks = c(12,13,14,15,16,17,18,19,20,21))
  
```
Plotting 100 student sample - based on color
```{r}
tiny_sample_lag <- 
  tiny_sample %>% 
  group_by(student_id_scram) %>% 
  mutate(n_age = lead(age)) %>%
  mutate(n_grade_level = lead(grade_level)) %>%
  filter(!is.na(n_grade_level),!is.na(n_age))

tiny_sample_lag %>%
  group_by(age,n_age,grade_level,n_grade_level,label) %>%
  summarize(Number_of_Students=n())%>%
  ggplot(aes(age,grade_level)) +
  geom_segment(aes(x = age, y = grade_level, xend = n_age, yend = n_grade_level,color=Number_of_Students))+
  scale_colour_gradient(low = 'red', high='blue', guide = "colourbar") + 
  scale_x_continuous(breaks = c(12,13,14,15,16,17,18,19,20,21)) +
  facet_wrap(~label)
```
Now based on thickness
```{r}
tiny_sample_lag %>%
  group_by(age,n_age,grade_level,n_grade_level,label) %>%
  summarize(Number_of_Students=n())%>%
  ggplot(aes(age,grade_level)) +
  geom_segment(aes(x = age, y = grade_level, xend = n_age, yend = n_grade_level,size=Number_of_Students))+
  scale_x_continuous(breaks = c(12,13,14,15,16,17,18,19,20,21)) +
  facet_wrap(~label)
```
##New Samples
Seperating by borough
```{r}
bios_by_borough <- bios %>%mutate(borough = substr(dbn,3,3))

set.seed(11)
borough_sample_name <- bios_by_borough %>% 
  select(student_id_scram,borough) %>%
  unique() %>%
  group_by(borough)%>%
  sample_n(100) %>%
  ungroup() %>%
  select(student_id_scram)
borough_sample <- left_join(borough_sample_name,bios_by_borough, by=c("student_id_scram"))

borough_lag <- 
  bios_sample %>%
  group_by(student_id_scram) %>% 
  mutate(n_age = lead(age)) %>%
  mutate(n_grade_level = lead(grade_level)) %>%
  filter(!is.na(n_grade_level),!is.na(n_age))

borough_lag %>%
  group_by(age,n_age,grade_level,n_grade_level,borough) %>%
  summarize(Number_of_Students=n())%>%
  ggplot(aes(age,grade_level)) +
  geom_segment(aes(x = age, y = grade_level, xend = n_age, yend = n_grade_level,color=Number_of_Students))+
  scale_colour_gradient(low = 'red', high='blue', guide = "colourbar") +
  facet_wrap(~borough)
```
Seperating by ethnicity
```{r}
bios_by_ethnicity <- bios
set.seed(11)
ethnicity_sample_name <- bios_by_ethnicity %>% 
  select(student_id_scram,ethnicity) %>% 
  unique() %>% 
  group_by(ethnicity)%>%
  sample_n(100)%>%
  ungroup()%>%
  select(student_id_scram)
ethnicity_sample <- left_join(ethnicity_sample_name,bios_by_ethnicity, by=c("student_id_scram"))

ethnicity_lag <- 
  ethnicity_sample %>% 
  group_by(student_id_scram) %>% 
  mutate(n_age = lead(age)) %>%
  mutate(n_grade_level = lead(grade_level)) %>%
  filter(!is.na(n_grade_level),!is.na(n_age))

ethnicity_lag %>%
  group_by(age,n_age,grade_level,n_grade_level,ethnicity) %>%
  summarize(Number_of_Students=n())%>%
  ggplot(aes(age,grade_level)) +
  geom_segment(aes(x = age, y = grade_level, xend = n_age, yend = n_grade_level,color=Number_of_Students))+
  scale_colour_gradient(low = 'red', high='blue', guide = "colourbar") +
  facet_wrap(~ethnicity)
```
Seperating by gender
```{r}
bios_by_gender <- bios
colnames(bios_by_gender)[4] <- "gender"

set.seed(11)
gender_sample_name <- bios_by_gender %>% 
  select(student_id_scram,gender) %>% 
  unique() %>% 
  group_by(gender) %>%
  sample_frac(0.0002) %>%
  ungroup() %>%
  select(student_id_scram)
gender_sample <- left_join(gender_sample_name,bios_by_gender, by=c("student_id_scram"))

gender_lag <- 
  gender_sample %>% 
  group_by(student_id_scram) %>% 
  mutate(n_age = lead(age)) %>%
  mutate(n_grade_level = lead(grade_level)) %>%
  filter(!is.na(n_grade_level),!is.na(n_age))

gender_lag %>%
  group_by(age,n_age,grade_level,n_grade_level,gender) %>%
  summarize(Number_of_Students=n())%>%
  ggplot(aes(age,grade_level)) +
  geom_segment(aes(x = age, y = grade_level, xend = n_age, yend = n_grade_level,color=Number_of_Students))+
  scale_colour_gradient(low = 'red', high='blue', guide = "colourbar") +
  facet_wrap(~gender)
```