---
title: "20170720_Student_Performance"
author: "Keri and David"
date: "7/20/2017"
output: html_document
---
Feature Taskforce for Student Performance keyed by student id and year
Relevant files from:
Science and Social Studies - science and social studies scores 
Math and ELA - math and ela scores
Regents - regents scores
Grad State Calc - states student's final outcome
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, root.dir = '/data/nycdoe/')
options(tibble.width=Inf)
library(ggplot2)
library(readr)
library(tidyverse)
library(reshape2)
library(stringr)
```
#Calling in Grad State Calc Data
Outcome_Category:
0 - Unknown, 1 - Dropped Out
2 - Still Enrolled, 3 - GED
4 - IEP, 5 - Local Diploma
6 - Regents, 7 - Advanced Regents
```{r}
read_grad_state_with_year <- function(filename) {
  year <- as.numeric(gsub('-.*$', '', basename(filename)))
  df <- read_csv(filename, col_types = cols(student_id_scram = col_character()))
  df$year <- year
  df <- select(df, year, student_id_scram, outcome_category)
}
files <- Sys.glob('/data/nycdoe/Grad/*.csv')
grad_state_backup <- map_df(files, read_grad_state_with_year)
grad_state <- grad_state_backup
```
Calling in ELA and Math Data
```{r}
read_ela_math_with_year <- function(filename) {
  year <- as.numeric(gsub('-.*$', '', basename(filename)))
  df <- read_csv(filename, col_types = cols(student_id_scram = col_character()))
  df$year <- year
  df <- select(df, year, student_id_scram, ela_test_grade, ela_perf_level, math_test_grade, math_perf_level)
}
files <- Sys.glob('/data/nycdoe/Math and ELA/*.csv')
ela_math_backup <- map_df(files, read_ela_math_with_year)
ela_math <- ela_math_backup
```
Calling in Science and Social Studies Data (2010-2015)
#Science (2005-2009) and 2010 - 2015 is a OK
```{r, echo=TRUE}
read_sci_with_year <- function(filename) {
  year <- as.numeric(gsub('-.*$', '', basename(filename)))
  df <- read_csv(filename, col_types = cols(student_id_scram = col_character(), grade = col_integer(), 
                                            performance_level = col_integer()))
  df$year <- year
  df <- select(df, year, student_id_scram, grade, subject, performance_level)
  }
files <- Sys.glob('/data/nycdoe/Science and Social Studies/*Science*.csv')
sci_backup <- map_df(files, read_sci_with_year)
sci <- sci_backup

read_soc_with_year <- function(filename) {
  year <- as.numeric(gsub('-.*$', '', basename(filename)))
  df <- read_csv(filename, col_types = cols(student_id_scram = col_character(),grade = col_integer(), 
                                            performance_level = col_integer()),
                 col_names = c("student_id_scram","dbn","grade","raw_score","scale_score","performance_level"))
  df$year <- year
  df$subject <- "Social Studies"
  df <- select(df, year, student_id_scram, grade, subject, performance_level)
  }
files <- Sys.glob('/data/nycdoe/Science and Social Studies/*SocStudies*.csv')
soc_backup <- map_df(files, read_soc_with_year)
soc <- soc_backup

#Combining Sci and Social Studies data
sci_soc_backup <- bind_rows(sci, soc)
sci_soc <- sci_soc_backup
```
Formatting  Sci Soc Data 
```{r}
sci_soc_performance <- sci_soc %>%
  select(-grade) %>%
  group_by(student_id_scram,year,subject) %>%
  summarize(performance_level=max(performance_level))%>%
  spread(key = subject, value = performance_level)

sci_soc_grade <- sci_soc %>%
  select(-performance_level) %>%
  group_by(student_id_scram,year,subject) %>%
  summarize(grade=max(grade))%>%
  spread(key = subject, value = grade)

colnames(sci_soc_performance) =  c("student_id_scram","year", "sci_perf_level", "soc_perf_level")
colnames(sci_soc_grade) =  c("student_id_scram","year", "sci_grade_level", "soc_grade_level")

sci_soc_backup <- full_join(sci_soc_performance,sci_soc_grade, by = c("student_id_scram","year"))
sci_soc <- sci_soc_backup
```
Calling in Regents Data
```{r}
read_regents_with_year <- function(filename) {
  year <- as.numeric(gsub('-.*$', '', basename(filename)))
  df <- read_csv(filename, col_types = cols(student_id_scram = col_character()))
  df$year <- year
  df <- select(df, year, student_id_scram, mark, exam_code)
}

setwd('/data/nycdoe/Regents/')
files <- Sys.glob('/data/nycdoe/Regents/*.csv')
regents_backup <- map_df(files, read_regents_with_year)
regents <- regents_backup

```
Formatting the Regents Data
```{r}
#Convert regent codes to first 3 substrings
regents <- regents %>% mutate(exam_code= substr(exam_code,1,2))
  
#Get a list of codes we want to keep. Anything taken by less than 10000 students is dropped
regents_keep<-regents %>%
  mutate(exam_code= substr(exam_code,1,2)) %>% 
  group_by(exam_code) %>% 
  count() %>%
  filter(n>10000) %>%
  select(exam_code)

#Bring regents down to the smaller list made earlier and then set all marks to the highest test score of one particular regent. Then spread so that a student has only one row
regents <- left_join(regents_keep,regents,by="exam_code")%>%
  group_by(student_id_scram, exam_code,year) %>%
  summarise(mark = max(mark))%>%
  spread(key=exam_code,value=mark)

#Change the names so that it can be bound to another dataframe
colnames(regents) =  c("student_id_scram","year", "Regent_English", "Regent_Language", "Regent_History", "Regent_math","Regent_RCT","Regent_Science")

```
Reading in Courses and Grades
```{r}
read_gpa_with_year <- function(filename) {
  year <- as.numeric(gsub('-.*$', '', basename(filename)))
  df <- read_csv(filename, col_types = cols_only(student_id_scram = col_character(), grade = col_integer(),markingperiod = col_integer(), 
                                                 numericequivalent = col_integer(), student_id = col_character(), coursetitle = col_character()))
  df$year <- year
  df <- select(df, year, student_id, grade, coursetitle, markingperiod, numericequivalent)
  }
files <- Sys.glob('/data/nycdoe/Courses and Grades/*.csv')
gpa_backup <- map_df(files[1:10], read_gpa_with_year)
gpa <- gpa_backup
```
formatting gpa
```{r}
gpa <- gpa %>%
  group_by(student_id,year) %>%
  summarize(GPA = mean(numericequivalent))
colnames(gpa) =  c("student_id_scram","year", "GPA")

```
Combining everything in one data frame 
```{r}
student_performance <- 
student_performance <- left_join(ela_math, grad_state, by = c("year","student_id_scram"))
student_performance <- left_join(student_performance, sci_soc, by = c("year","student_id_scram"))
student_performance <- left_join(student_performance, regents, by = c("year","student_id_scram"))
student_performance <- left_join(student_performance, gpa, by = c("year","student_id_scram"))
```
Saving everything
```{r}
save(student_performance, file = "/data/nycdoe/clean_data/student_perf_feature.Rdata")
```