---
title: '20170718'
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Loading in data.
```{r}
library(tidyverse)
library(ggplot2)
library(scales)
library(directlabels)

setwd('/data/nycdoe/')
files <- Sys.glob('/data/nycdoe/June Biog/*.csv')

read_csv_with_year <- function(filename) {
  year <- as.numeric(gsub('-.*$', '', basename(filename)))
  df <- read_csv(filename, col_types = cols(grade_level=col_character(), grade_code=col_character(), admit_code=col_character(), adcode=col_character(),fall_days_present = col_character(), spring_days_present = col_character()))
  df$year <- year
  df
}

bios <- map_df(files, read_csv_with_year) %>%
  filter(grade_level!="AD", grade_level!="IN") %>%
  mutate(grade_level=ifelse(grade_level=="0K", 0, 
                            ifelse(grade_level=="PK", -1, as.numeric(grade_level)))) 
```
# 20170717
## Per-grade "survival"
```{r}
(df <- 
    bios %>%
    #filter(student_id_scram==420002000) %>%
    #arrange(student_id_scram) %>%
    select(student_id_scram, year, grade_level) %>%
    mutate(grade_level=as.numeric(grade_level)) %>%
    group_by(student_id_scram) %>%
    arrange(year) %>%
    mutate(did_continue=ifelse(!is.na(lead(grade_level)) & grade_level == lead(grade_level) - 1, 1, 0)) )
  
exit_rates <- df %>%
    group_by(year, grade_level) %>%
    #distinct(year, student_id_scram, did_continue) %>%
    #group_by(student_id_scram, grade_level) %>%
    summarise(continuing_students=sum(did_continue), total_students=n(), exit_rate=1-continuing_students/total_students)
```

6 years with the lowest exit rates
```{r}
exit_rates %>%
  arrange(exit_rate) %>%
  head()
```

6 years with the highest exit rates
```{r}
exit_rates %>%
  arrange(desc(exit_rate)) %>%
  filter(grade_level!=12, year!=2015) %>%
  head()
```

##Per-grade "survival" by school
```{r}
(df_with_schools <- 
    bios %>%
    #filter(student_id_scram==420002000) %>%
    #arrange(student_id_scram) %>%
    select(student_id_scram, year, grade_level, dbn) %>%
    mutate(grade_level=as.numeric(grade_level)) %>%
    group_by(student_id_scram) %>%
    arrange(year) %>%
    mutate(did_continue=ifelse(!is.na(lead(grade_level)) & grade_level == lead(grade_level) - 1, 1, 0)) )
  
exit_rates_by_school <- 
  df_with_schools %>%
    group_by(dbn, year, grade_level) %>%
    #distinct(year, student_id_scram, did_continue) %>%
    #group_by(student_id_scram, grade_level) %>%
    summarise(continuing_students=sum(did_continue), total_students=n(), exit_rate=1-continuing_students/total_students)
```

6 schools with the lowest exit rates per year/grade
```{r}
exit_rates_by_school %>%
  arrange(exit_rate) %>%
  head()
```

6 schools with the highest exit rates per year/grade
```{r}
exit_rates_by_school %>%
  arrange(desc(exit_rate)) %>%
  filter(grade_level!=12, year!=2015) %>%
  head()
```

6 schools with the lowest entrance rates
```{r}
df_with_schools %>%
  group_by(dbn) %>%
  #distinct(year, student_id_scram, did_continue) %>%
  #group_by(student_id_scram, grade_level) %>%
  summarise(continuing_students=sum(did_continue), total_students=n(), exit_rate=1-continuing_students/total_students) %>%
  arrange(exit_rate) %>%
  head()
```

6 schools with the highest entrance rates
```{r}
df_with_schools %>%
  group_by(dbn) %>%
  #distinct(year, student_id_scram, did_continue) %>%
  #group_by(student_id_scram, grade_level) %>%
  filter(grade_level!=12, year!=2015) %>%
  summarise(continuing_students=sum(did_continue), total_students=n(), exit_rate=1-continuing_students/total_students) %>%
  arrange(desc(exit_rate)) %>%
  head()
```

## How many students enter in each grade
```{r}
(df_entered <- 
    bios %>%
    #filter(student_id_scram==420002000) %>%
    #arrange(student_id_scram) %>%
    select(student_id_scram, year, grade_level) %>%
    mutate(grade_level=as.numeric(grade_level)) %>%
    group_by(student_id_scram) %>%
    arrange(year) %>%
    mutate(did_enter=ifelse(is.na(lag(grade_level)) | grade_level != lag(grade_level) + 1, 1, 0)) )
  
entrance_rates <- 
  df_entered %>%
    group_by(year, grade_level) %>%
    #distinct(year, student_id_scram, did_continue) %>%
    #group_by(student_id_scram, grade_level) %>%
    summarise(entering_students=sum(did_enter), total_students=n(), entrance_rate=entering_students/total_students)

entrance_rates
```

6 years with the lowest entrance rates
```{r}
entrance_rates %>%
  arrange(entrance_rate) %>%
  head()
```

6 years with the highest entrance rates (besides 1st grade)
```{r}
entrance_rates %>%
  arrange(desc(entrance_rate)) %>%
  filter(grade_level!=-1, year!=2005) %>%
  head()
```

##How many students enter in each grade by school
```{r}
(df_entered_with_schools <- 
    bios %>%
    #filter(student_id_scram==420002000) %>%
    #arrange(student_id_scram) %>%
    select(student_id_scram, year, grade_level, dbn) %>%
    mutate(grade_level=as.numeric(grade_level)) %>%
    group_by(student_id_scram) %>%
    arrange(year) %>%
    mutate(did_enter=ifelse(is.na(lag(grade_level)) | grade_level != lag(grade_level) + 1, 1, 0)) )
  
entrance_rates_by_school <- 
  df_entered_with_schools %>%
    group_by(dbn, year, grade_level) %>%
    #distinct(year, student_id_scram, did_continue) %>%
    #group_by(student_id_scram, grade_level) %>%
    summarise(entering_students=sum(did_enter), total_students=n(), entrance_rate=entering_students/total_students)

entrance_rates_by_school
```

6 schools with the lowest entrance rates by year/grade
```{r}
entrance_rates_by_school %>%
  arrange(entrance_rate) %>%
  head()
```

6 schools with the highest entrance rates by year/grade (except 1st grade)
```{r}
entrance_rates_by_school %>%
  arrange(desc(entrance_rate)) %>%
  filter(grade_level!=-1, year!=2005) %>%
  head()
```

6 schools with the lowest entrance rates
```{r}
df_entered_with_schools %>%
  group_by(dbn) %>%
  #distinct(year, student_id_scram, did_continue) %>%
  #group_by(student_id_scram, grade_level) %>%
  summarise(entering_students=sum(did_enter), total_students=n(), entrance_rate=entering_students/total_students)  %>%
  arrange(entrance_rate) %>%
  head()
```

6 schools with the highest entrance rates
```{r}
df_entered_with_schools %>%
  group_by(dbn) %>%
  #distinct(year, student_id_scram, did_continue) %>%
  #group_by(student_id_scram, grade_level) %>%
  filter(grade_level!=-1, year!=2005) %>%
  summarise(entering_students=sum(did_enter), total_students=n(), entrance_rate=entering_students/total_students)  %>%
  arrange(desc(entrance_rate)) %>%
  head()
```

## "True survivors" - what fraction of students who started in grade K graduate high school 12 years later?

```{r}
(entries_per_student <-
  df %>%
  group_by(student_id_scram) %>%
  summarise(count=n()))

entries_per_student %>%
  group_by(count) %>%
  count()
```

Just for high school.

Question - since the data set ends at 2015, the 9th, 10th, and 11th graders are considered as only staying 1 year, even though they might have stayed longer. Filtering out everything after 2012 (the last year for which we have data for each student for 4 years from that point) doesn't solve the problem. How can we stagger the filters, so that in 2015 we only take the 12th graders, 2014 we take 11th and 12th, and 2013 we take 10th 11th and 12th?
```{r}
# (entries_per_student <-
#   df %>%
#   group_by(student_id_scram) %>%
#   filter(grade_level>8) )

(entries_per_student <-
    df %>%
    group_by(student_id_scram) %>%
    filter(ifelse(year==2013, grade_level>9, 
                  ifelse(year==2014, grade_level>10,
                         ifelse(year==2015, grade_level>11, grade_level>8)))))

# (entries_per_student <-
#   entries_per_student %>%
#   filter(grade_level>8) %>%
#   #filter(year<2013) %>%
#   summarise(count=sum(did_continue)) )

entries_per_student %>%
  summarise(count=sum(did_continue)) %>%
  summarise(num_survivors=sum(count==3), perc_survivors=num_survivors/n())
```

#20170718

## Visualizations

```{r}
continuation_rates <- df %>%
    group_by(year, grade_level) %>%
    #distinct(year, student_id_scram, did_continue) %>%
    #group_by(student_id_scram, grade_level) %>%
    summarise(continuing_students=sum(did_continue), total_students=n(), continuation_rate=continuing_students/total_students)

continuation_rates %>%
  group_by(grade_level) %>%
  summarise(avg_continuation=mean(continuation_rate), se=sd(continuation_rate)/n()) %>%
  ggplot(aes(x=grade_level, y=avg_continuation)) +
  scale_x_continuous(breaks = seq(-2, 12, 1)) +
  scale_y_continuous(labels = percent) +
  geom_point() +
  geom_line() +
  labs(x="Grade Level", y="Average Percent Continuation", title="Year-to-Year Continuation") +
  geom_errorbar(aes(ymin = avg_continuation-se, ymax = avg_continuation+se))
```

```{r}
continuation_rates %>%
  filter(grade_level!=12, year!=2015) %>%
  ggplot(aes(x=year, y=continuation_rate, color=as.factor(grade_level))) +
  geom_point() +
  geom_line() +
  labs(x="Academic Year", y="Percent Continuation", title="Trends in Continuation Rates by Grade Level") +
  scale_y_continuous(labels = percent) +
  scale_x_continuous(breaks=seq(2005, 2014, 1), expand=c(0, 1)) +
  geom_dl(aes(label = as.factor(grade_level)), method = list(dl.trans(x = x + .3), "last.points", cex = 1))
```

```{r}
new_df <-
  entries_per_student %>%
  #group_by(student_id_scram) %>%
  filter(any(grade_level==9)) %>%
  group_by(year) %>%
  #distinct(student_id_scram) %>%
  summarise(num_ninth=sum(grade_level==9), 
            continued_to_tenth=sum(grade_level==9 & did_continue==1, na.rm=T), 
            continued_to_eleventh=sum(grade_level==10 & did_continue==1, na.rm=T), 
            continued_to_twelfth=sum(grade_level==11 & did_continue==1, na.rm=T),
            perc_ninth=num_ninth/num_ninth,
            perc_tenth=continued_to_tenth/num_ninth,
            perc_eleventh=continued_to_eleventh/num_ninth,
            perc_twelfth=continued_to_twelfth/num_ninth) 

new_df

new_df_2 <-
  new_df %>%
  select(year, 6:9) %>%
  gather("grade", "num_students", 2:5) %>%
  arrange(year)

new_df_2$grade[new_df_2$grade=="perc_ninth"] <- 9
new_df_2$grade[new_df_2$grade=="perc_tenth"] <- 10
new_df_2$grade[new_df_2$grade=="perc_eleventh"] <- 11
new_df_2$grade[new_df_2$grade=="perc_twelfth"] <- 12

new_df_2 <-
  new_df_2 %>%
  mutate(grade=as.numeric(grade))

new_df_2

new_df_2 %>%
  filter(year<=2012, year>2006) %>%
  ggplot(aes(x=grade, y=num_students, color=as.factor(year))) +
  geom_point() +
  geom_line() +
  scale_y_continuous(labels = percent) +
  scale_x_continuous(breaks=seq(9, 12, 1), expand=c(0, 1)) +
  geom_dl(aes(label = as.factor(year)), method = list(dl.trans(x = x + .3), "last.points", cex = 0.8)) +
  labs(x="Grade Level", y="Percent Remaining", title="Student Drop-Outs Throughout High School", color="Class of")
```

The general trends are consistent with the data found on nyc.gov, but our values are about 10% lower. Maybe because this graph is showing us what percent of the students made it to 12th grade, but it's not showing us how many of those 12th graders graduate. How can we add in, for each student, whether they graduated?

We had to drop years after 2012 - 2012 is the last year that we have data for those 9th graders through all 4 years of college.

We had to drop 2005 and 2006 because for some reason, it was telling us that in 2005 almost everyone dropped out after 10th grade and in 2006 almost everyone dropped out after 11th grade.