---
title: "Ethnicity by school"
author: "Rivka"
date: "7/25/2017"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(ggplot2)
library(readr)
library(tidyverse)
library(dplyr)

#load june_bio (nothing has been done to this df--not cleaned)

load("/data/nycdoe/clean_data/bio_data.Rdata")


```
clean bio data (remove na student_ids)
```{r}
bio_data <- bio_data %>% filter(!is.na(student_id_scram))

```

Get number students in each school across all years
```{r}

bio_data %>% group_by(dbn) %>% summarise(totalNumStudents = n())

```
Plot number in each school ##ERROR
```{r}
bio_data %>% 
  group_by(dbn, student_id_scram) %>% 
  ggplot(aes(x = dbn)) + geom_histogram(stat = count)

```

number schools in our bio data file--2,312 (!!!!!!!!!!!!!!)
```{r}
nrow(bio_data %>% group_by(dbn) %>% 
  summarise())

```

#ERROR I think 0 included everything...staggering in weird way
plot number of schools with same num students
```{r}
bio_data %>% group_by(dbn) %>% 
  summarise(totalNumStudents = n()) %>% 
  ggplot(aes(x = totalNumStudents)) + 
  geom_histogram() +
  ylab("Num schools")
```

plot schools with same number students facet by year
```{r}
bio_data %>% group_by(dbn, year) %>% 
  summarise(totalNumStudents = n()) %>% 
  ggplot(aes(x = totalNumStudents)) + 
  geom_histogram() +
  ylab("Num schools") + 
  facet_wrap(~year)

```

plot number in each school facet by year
```{r}
bio_data %>% group_by(dbn, year) %>% 
  summarise(totalNumStudents = n()) %>% 
  ggplot(aes(y = totalNumStudents, x = dbn)) + 
  geom_bar(stat = "identity") +
  ylab("Num schools") + 
  facet_wrap(~year)

```

count number of students each ethnicity over all years 
```{r}
bio_data %>% group_by(student_id_scram, ethnicity) %>% summarise() %>% group_by(ethnicity) %>% summarise(count = n())
```
count ethnicity in each school
```{r}
#count Asian
asian_count <- bio_data %>% group_by(dbn, ethnicity == 'Asian', year)%>% summarise(count = n()) %>% filter(`ethnicity == "Asian"` == TRUE)

#count Black
black_count <- bio_data %>% group_by(dbn, ethnicity == 'Black', year)%>% summarise(count = n()) %>% filter(`ethnicity == "Black"` == TRUE)

#count Hispanic
hispanic_count <- bio_data %>% group_by(dbn, ethnicity == 'Hispanic', year)%>% summarise(count = n()) %>% filter(`ethnicity == "Hispanic"` == TRUE)

#count native american
native_american_count <- bio_data %>% group_by(dbn, ethnicity == 'Native American', year)%>% summarise(count = n()) %>% filter(`ethnicity == "Native American"` == TRUE)

#white
white_count <- bio_data %>% group_by(dbn, ethnicity == 'White', year)%>% summarise(count = n()) %>% filter(`ethnicity == "White"` == TRUE)

#multi-racial
multi_count <- bio_data %>% group_by(dbn, ethnicity == 'Multi-Racial', year)%>% summarise(count = n()) %>% filter(`ethnicity == "Multi-Racial"` == TRUE)

```
number students in each dbn each year
```{R}
num_students_dbn <- bio_data %>% group_by(dbn, year) %>% summarise(totalNumStudents = n())
```

join the race counts with student count by dbn and year
```{r}
all_counts <- inner_join(asian_count, hispanic_count, by = c("dbn","year"))

all_counts <- inner_join(all_counts, white_count, by = c("dbn", "year"))

all_counts <- inner_join(all_counts, black_count, by = c("dbn", "year"))
all_counts <- inner_join(all_counts, native_american_count, by = c("dbn", "year"))
all_counts <- inner_join(all_counts, multi_count, by = c("dbn", "year"))

#rename cols
colnames(all_counts)[4] <- "asian_count"
colnames(all_counts)[6] <- "hispanic_count"
colnames(all_counts)[8] <- "white_count"
colnames(all_counts)[10] <- "black_count"
colnames(all_counts)[12] <- "native_amer_count"
colnames(all_counts)[14] <- "multi_count"

all_counts <- all_counts %>% ungroup() %>% select(c(1,3,4,6,8,10,12,14))

all_counts <- inner_join(num_students_dbn, all_counts, by = c("dbn", "year"))

```
Get percents for those values
```{r}
all_counts <- all_counts %>% mutate(percentAsian = asian_count/totalNumStudents *100, 
                                    percentBlack = black_count/totalNumStudents*100, 
                                    percentWhite = white_count/totalNumStudents*100, 
                                    percentHispanic = hispanic_count/totalNumStudents*100, 
                                    percentNativeAmer = native_amer_count/totalNumStudents*100, 
                                    percentMulti =  multi_count/totalNumStudents*100)

```

