---
title: "20170726 - Do You Get What You Want (cont. 2)"
author: "Anandini, Thoa"
date: "7/26/2017"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
#setwd('/data/nycdoe/')
hsaps_files <- Sys.glob('/data/nycdoe/HSAPS/*.csv')
june_biog_files <- Sys.glob('/data/nycdoe/June Biog/*.csv')
```
## Read in data from HSAPS

Here we read in HSAPS data from 2009 to 2015 (in academic year term: 2009-10 to 2015-16) because they have small proportion of NAs in `final_disposition`, and they have `lag_xxx_applied`.

We choose the following colums:

* `r1programcode1` : specify the program that a student ranks as 1st choice in the first round.
* `r1programtype1`: specify the most likely admissions method used by the high school program applied to by the student 
* `r1eligibility1`: specify whether a student is eligible to apply to the selected program. Note: it is possible that the student does not fit any of the program's priority groups, and, therefore, is not eligible for admission.
* `final_disposition`: specify the program that a student gets finalized to.
* `shsat_tested`: specify whether a student took a Specialized HS test
* `shsat_offered`: specify whether a student got an offer from a specialized HS.
* `any_log_offer`: specify whether a student got an offer from LaGuardia HS.
* `lag_xxx_applied`: specify whether a student applied to the program xxx at LaGuardia.
* `lag_xxx_offer`: specify whether a student got an offer from the program xxx at LaGuardia.
* `r1programcode2`and `r1programcode3`: this is for "school combo" network

```{r read-hsaps, results='hide', message=TRUE}
# read hsaps data
read_satisfied_w_lag_3ranks <- function(filename) {
  print(filename)
  year <- as.numeric(gsub('-.*$', '', basename(filename)))
  df <- read_csv(filename, col_types = cols(student_id_scram=col_character(),
                                            r1eligibility1=col_character(),
                                            final_disposition_round=col_character(),
                                            shsat_offer=col_character(),
                                            shsat_tested=col_character(),
                                            lag_arts_offer=col_character(),
                                            lag_arts_applied=col_character(),
                                            lag_dance_offer=col_character(),
                                            lag_dance_applied=col_character(),
                                            lag_instr_offer=col_character(),
                                            lag_instr_applied=col_character(),
                                            lag_vocal_offer=col_character(),
                                            lag_vocal_applied=col_character(),
                                            lag_drama_offer=col_character(),
                                            lag_drama_applied=col_character(),
                                            lag_theatre_offer=col_character(),
                                            lag_theatre_applied=col_character()))
  df$year <- year
  df <- df %>% 
    select(year, 
           student_id_scram, 
           r1programcode1,
           r1programtype1,
           r1eligibility1,
           r1programcode2,
           r1programcode3,
           contains("final_disposition"),
           shsat_tested,
           shsat_offer,
           any_lag_offer,
           starts_with("lag"),
           borough,
           feeder_borough,
           feeder_dbn)
}

satisfied_09_15 <- map_df(hsaps_files[5:11], read_satisfied_w_lag_3ranks)
```

## Clean HSAPS features

```{r clean}

# reformat admission method
satisfied_09_15 <-
  satisfied_09_15 %>%
  mutate(r1programtype1 = ifelse(r1programtype1 == "Ed. opt.",
                                  yes = "Ed. Opt.",
                                  no = r1programtype1))
# reformat student's eligibility: NA also means NA in r1programcode1, so we will set that to 0.
satisfied_09_15 %>% filter(is.na(r1programcode1)) %>% count(r1eligibility1, year)
satisfied_09_15$r1eligibility1[is.na(satisfied_09_15$r1eligibility1)] <- "0"
satisfied_09_15 <-
  satisfied_09_15 %>%
  mutate(r1eligibility1 = ifelse(r1eligibility1 %in% c("0", "N"),
                                 yes = 0,
                                 no = 1))


# drop 115 students who are NOT from 1 of the 5 boroughs: 572,603 rows become 572,488 rows
satisfied_09_15 <-
  satisfied_09_15 %>%
  filter(borough %in% c("M", "K", "Q", "R", "X"))

#### SHSAT_TESTED #### 
# because we don't know what NA means, and since they account for a relatively small amount, we will covert it to 0 (just like N).
# covers 3 cases: (NA is seperately treated first)
# 0 or 1 ==> stay same
# Y  ==> 1
# N ==> 0
satisfied_09_15$shsat_tested[is.na(satisfied_09_15$shsat_tested)] <- "0"
satisfied_09_15 <-
  satisfied_09_15 %>%
  mutate(shsat_tested = ifelse(shsat_tested %in% c("0", "1"),
                              yes = shsat_tested,
                              no = ifelse(shsat_tested == "Y",
                                          yes = "1",
                                          no = "0")))
satisfied_09_15 <-
  satisfied_09_15 %>%
  mutate(shsat_tested = as.numeric(shsat_tested))


#### SHSAT_OFFER ####
satisfied_09_15$shsat_offer[is.na(satisfied_09_15$shsat_offer)] <- "0"
satisfied_09_15 %>% count(shsat_offer)
satisfied_09_15 <-
  satisfied_09_15 %>%
  mutate(shsat_offer = ifelse(shsat_offer == "0",
                              yes = "0",
                              no = "1"))  
satisfied_09_15 <-
  satisfied_09_15 %>%
  mutate(shsat_offer = as.numeric(shsat_offer))

#### ANY_LAG_OFFER ####
# since we're only interested in whether or not you get offered by LaGuardia, we will create a new binary variable for this.
satisfied_09_15 <-
  satisfied_09_15 %>%
  mutate(lag_offered = as.numeric(any_lag_offer > 0))

#### LAGUARDIA APPLIED ######
# the NA's in 2014 and 2015 are so consistent. When we checked, they do not have r1programcode1-2-3, type1, or eligibility1. One third of them did NOT take the SHSAT, while two-thirds have NA so we can't know. 1,467 of them took shsat. shsat_offer is all NA.
satisfied_09_15 %>%
  filter(is.na(lag_arts_applied)) #%>% count(shsat_tested)

# as said before, we will drop applicants who (1) not participate in any kind of specialized application AND (2) not have r1programcode1. That means we will drop these 3722 students. 574,488 rows go down to 568,766 rows.
satisfied_09_15 <-
  satisfied_09_15 %>%
  filter(!is.na(lag_arts_applied))

# now, covert the formats and types for binary values.

## examine current value types:
satisfied_09_15 %>% count(lag_arts_applied)
satisfied_09_15 %>% count(lag_arts_offer)


## convert:
satisfied_09_15$lag_arts_offer[is.na(satisfied_09_15$lag_arts_offer)] <- "0"
satisfied_09_15$lag_dance_offer[is.na(satisfied_09_15$lag_dance_offer)] <- "0"
satisfied_09_15$lag_instr_offer[is.na(satisfied_09_15$lag_instr_offer)] <- "0"
satisfied_09_15$lag_vocal_offer[is.na(satisfied_09_15$lag_vocal_offer)] <- "0"
satisfied_09_15$lag_drama_offer[is.na(satisfied_09_15$lag_drama_offer)] <- "0"
satisfied_09_15$lag_theatre_offer[is.na(satisfied_09_15$lag_theatre_offer)] <- "0"
satisfied_09_15 <-
  satisfied_09_15 %>%
  mutate(lag_arts_applied = ifelse(lag_arts_applied %in% c("Y", "1"), 1, 0),
         lag_dance_applied = ifelse(lag_dance_applied %in% c("Y", "1"), 1, 0),
         lag_instr_applied = ifelse(lag_instr_applied %in% c("Y", "1"), 1, 0),
         lag_vocal_applied = ifelse(lag_vocal_applied %in% c("Y", "1"), 1, 0),
         lag_drama_applied = ifelse(lag_drama_applied %in% c("Y", "1"), 1, 0),
         lag_theatre_applied = ifelse(lag_theatre_applied %in% c("Y", "1"), 1, 0),
         lag_arts_offer = ifelse(lag_arts_offer %in% c("OFFER", "1"), 1, 0),
         lag_dance_offer = ifelse(lag_dance_offer %in% c("OFFER", "1"), 1, 0),
         lag_instr_offer = ifelse(lag_instr_offer %in% c("OFFER", "1"), 1, 0),
         lag_vocal_offer = ifelse(lag_vocal_offer %in% c("OFFER", "1"), 1, 0),
         lag_drama_offer = ifelse(lag_drama_offer %in% c("OFFER", "1"), 1, 0),
         lag_theatre_offer = ifelse(lag_theatre_offer %in% c("OFFER", "1"), 1, 0))
```

## Calculate the `was_satisfied` metrics

```{r calculate-was-satisfied}

# drop 16,762 students who NOT participated in any of these: (1) listing a top1 choice, (2) took shsat, (3) applied to Laguardia. 568,766 rows go down to 552,004 rows. (In the "Clean" chunk above, we removed rows with NA's in lag_applied that are unclear elsewhere. Here, the drop is stronger because we clear up "NA in top1 AND 0 in test AND 0 in applied")
satisfied_09_15 <-  
  satisfied_09_15 %>%
  filter(!is.na(r1programcode1) |
          shsat_tested == 1 |
          lag_arts_applied == 1 | 
          lag_dance_applied == 1 | 
          lag_instr_applied == 1 | 
          lag_vocal_applied == 1 | 
          lag_drama_applied == 1 | 
          lag_theatre_applied == 1)

# # those applied to Arts at LaGuardia do NOT have final_disposition matching top1 choice. This means the only way for Arts applicants to be satisfied is to get an offer from LaGuardia or SHSAT.
# temp %>%
#   filter(lag_arts_applied == 1 &  final_disposition_round==r1programcode1)

satisfied_09_15 <-
  satisfied_09_15 %>%
  mutate(was_satisfied = as.numeric((lag_arts_applied==1 & lag_arts_offer==1) |
                                    (lag_arts_applied==1 & final_disposition==r1programcode1) |
                                      
                                      (lag_dance_applied==1 & lag_dance_offer==1) |
                                      (lag_dance_applied==1 & final_disposition==r1programcode1) |
                                      
                                      (lag_instr_applied==1 & lag_instr_offer==1) |
                                      (lag_instr_applied==1 & final_disposition==r1programcode1) |
                                      
                                      (lag_vocal_applied==1 & lag_vocal_offer==1) |
                                      (lag_vocal_applied==1 & final_disposition==r1programcode1) |
                                      
                                      (lag_drama_applied==1 & lag_drama_offer==1) |
                                      (lag_drama_applied==1 & final_disposition==r1programcode1) |
                                      
                                      (lag_theatre_applied==1 & lag_theatre_offer==1) |
                                      (lag_theatre_applied==1 & final_disposition==r1programcode1) |
                                      
                                      (shsat_tested==1 & shsat_offer==1) |
                                      (shsat_tested==1 & final_disposition==r1programcode1) |
                                      
                                      (final_disposition==r1programcode1)
                                    ) 
         ) %>%
  replace_na(list(was_satisfied = 0))
```

## Load in additional data

#### Student's features (demo & perf) 

```{r load-student-perf}

# load in data on students' quantiled GPA
load('/data/nycdoe/clean_data/stud_perf_quantiledGPA.Rdata')
perstudent_perf <- stud_perf_quantGPA
rm(stud_perf_quantGPA)

# this table has 4.4 million rows
(perstudent_gpa_long <-
  perstudent_perf %>%
  filter(student_id_scram %in% satisfied_09_15$student_id_scram) %>%
  select(year, student_id_scram, grade_level, GPA, quantiled_GPA) %>%
  arrange(student_id_scram, year))
rm(perstudent_perf)

# NOT DONE: SPREAD table to have student quantGPA across years (i.e. 1 row per student_id), not so possible because the education path of students are not uniform
```

```{r load-student-demo, results='hide', message=TRUE}

# load in students' demographics from June Biog
read_students_hs <- function(filename) {
  year <- as.numeric(gsub('-.*$', '', basename(filename)))
  df <- read_csv(filename, col_types = cols(student_id_scram=col_character(),
                                            grade_level=col_integer()))
  df$year <- year
  df <- df %>% 
    select(year, student_id_scram, grade_level, "feeder_dbn(june)" = dbn, sex, ethnicity, ell, swd, home_lang, poverty)
}
students_hs_09_15 <- map_df(june_biog_files[5:11], read_students_hs)
```

#### School's features 

```{r load-new-comprehensive-school-features}

# load in NEW school features:
load('/data/nycdoe/clean_data/will_i_get_in_school_features.Rdata')
summary(will_i_get_in_school_features)
hs_features_new <-
  will_i_get_in_school_features %>%
  select(dbn, 
         "total_stud" = total_students,
         #"total_grad" = grads,
         "perc_grad" = percentGrad,
         #"total_poverty" = sumPoverty,
         "perc_poverty" = percentPoverty,
         #"total_White" = totalWhite,
         "perc_White" = percentWhite,
         #"total_Black" = totalBlack,
         "perc_Black" = percentBlack,
         #"total_Hispanic" = totalHispanic,
         "perc_Hispanic" = percentHispanic,
         #"total_MultiRacial" = totalMultiRacial,
         "perc_MultiRacial" = percentMulti,
         #"total_NativeAmer" = totalNativeAmer,
         "perc_NativeAmer" = percentNativeAmer,
         #"total_Asian" = totalAsian,
         "perc_Asian" = percentAsian,
         "attendance" = avgAtt,
         English,
         Math,
         History,
         Science)
rm(will_i_get_in_school_features)
names(hs_features_new)[12:15] <- paste(names(hs_features_new)[12:15], "regents", sep="_")
```

#### Programcode look-up

```{r load-programcode-dict}

# load in data that maps programcodes to schoolcodes - use this to map `final_disposition` to schools, in order to add school features into the model:
# 778 program codes in total.
load('/data/nycdoe/clean_data/schoolcodemapping.Rdata')
percode_schoolmapping <- minimal_allschoolcodes
rm(minimal_allschoolcodes)

##### CAN WE EXTRACT MORE CODE-SCHOOL ENTRIES? ########### (end result is 817 schools, an increase of 39 entries)

# there are 56 dbns (final_disposition_dbn) not existing in `percode_schoolmapping` (with one case being NA)
satisfied_09_15 %>%
  distinct(final_disposition_dbn) %>%
  filter(!(final_disposition_dbn) %in% percode_schoolmapping$dbn) 
# let's see if these missing dbn's have associated programcodes (i.e. final_disposition). If positive, we can create additional code-school mapping to add to percode_schoolmapping.
satisfied_09_15 %>%
  distinct(final_disposition_dbn, .keep_all = TRUE) %>%
  filter(!(final_disposition_dbn) %in% percode_schoolmapping$dbn) %>%
  filter(!is.na(final_disposition_dbn)) %>%
  group_by(final_disposition_dbn) %>%
  summarize(num_programs = length(unique(final_disposition))) %>% arrange(desc(num_programs))
# each of these dbn's have exactly one code mapped to it, hence we will now create an additional code-school mapping
(addl_code_school_mapping <-
    satisfied_09_15 %>%
    distinct(final_disposition_dbn, .keep_all = TRUE) %>%
    filter(!(final_disposition_dbn) %in% percode_schoolmapping$dbn) %>%
    filter(!is.na(final_disposition_dbn)) %>%
    select("dbn" = final_disposition_dbn,
           "pgcode" = final_disposition))
# do these 55 dbn's include Specialized HS? YES! 
addl_code_school_mapping %>%
  filter(substring(dbn, 3) %in% c("X445", "K449", "K430", "R605", "M475", "M692", "X696", "Q687", "M485"))
# but this does not seem complete, because LaGuardia should have 6 programs, so total number of rows must be 14, not 9. Hence we will remove these rows. Now rows go down to 46.
addl_code_school_mapping <-
  addl_code_school_mapping %>%
  filter(!(substring(dbn, 3) %in% c("X445", "K449", "K430", "R605", "M475", "M692", "X696", "Q687", "M485")))
# create a fuller code-school map to later fill in the gaps in missing r1r1-schools
# this fuller code-school map will help fill in r1r1_dbn for 2,242 applications
# withfeatures_satisfied_09_15 %>%
#   filter((r1programcode1 %in% addl_code_school_mapping$pgcode) & is.na(r1r1_dbn))# %>% count(is.na(r1r1_attendance))
addl_code_school_mapping <-
  addl_code_school_mapping %>%
  mutate(pgname = rep("Missing", times=nrow(addl_code_school_mapping)),
         school_name = rep("Missing", times=nrow(addl_code_school_mapping)),
         interest = rep("Missing", times=nrow(addl_code_school_mapping)),
         method = rep("Missing", times=nrow(addl_code_school_mapping)))
oldmap <- percode_schoolmapping %>%
  select(-year) %>%
  select(dbn, pgcode, everything()) %>% ungroup()
# it's important to make sure that when we bind 2 tables, they do not duplicate on pgcodes and give each pgcode >1 school. Hence, here we remove from the additional table whatever codes that already exist in the original mapping.
addl_code_school_mapping <-
  addl_code_school_mapping %>%
  filter(!(pgcode %in% oldmap$pgcode))
percode_schoolmapping_fuller <- rbind(oldmap, addl_code_school_mapping)
rm(addl_code_school_mapping, oldmap, percode_schoolmapping)
save(percode_schoolmapping_fuller, file = '/data/nycdoe/clean_data/schoolcodemapping_fuller.Rdata')
```

## Add features to main table

#### Student's features (demo & perf) 

```{r add-student-demo}

rm(withfeatures_satisfied_09_15)

# join applicants data with student's demographics: 552,044 rows go down to 485,824 rows because we restrict to (1) applicants must be 8th-graders in the year they apply, and (2) their feeder_dbn must match the 8th-grade dbn in June Biog.
withfeatures_satisfied_09_15 <-
  left_join(satisfied_09_15, students_hs_09_15, by=c("student_id_scram", "year")) %>%
  filter(grade_level == 8 & feeder_dbn == `feeder_dbn(june)`)
summary(withfeatures_satisfied_09_15)

# REMOVING DUPLICATE ENTRIES (keep the earliest year): NEED TO DISREGARD THE `YEAR` COLUMN. ONCE DONE, WILL DROP 170 ROWS
# withfeatures_satisfied_09_15 <-
#   withfeatures_satisfied_09_15 %>%
#   distinct(.keep_all = TRUE)
```

```{r add-student-pef}

# join applicants data with student-performance-quantGPA data. NOTE: if we join by year like this, we will lose performance in previous years! THIS ONLY ALLOWS 8th-GRADE (or 7th-GRADE) GPA.
perstudent_gpa_long <-
  perstudent_gpa_long %>%
  filter(student_id_scram %in% withfeatures_satisfied_09_15$student_id_scram)
withfeatures_satisfied_09_15 <-
  left_join(withfeatures_satisfied_09_15, perstudent_gpa_long, by=c("student_id_scram", "year")) %>%
  select(-grade_level.y) %>%  
  rename("grade_level" = grade_level.x)
summary(withfeatures_satisfied_09_15)
```

#### Middle school's features 

```{r add-middle-school-features}

# # yay! feeder_dbn is not NA at all!
# withfeatures_satisfied_09_15 %>%
#   count(is.na(feeder_dbn))

# join in MIDDLE school's features
middle_school_temp <- hs_features_new %>%
  select(-(12:15), -perc_grad)  # remove columns that do not apply to middle schools
names(middle_school_temp) <- paste("feeder", names(middle_school_temp), sep="_")
summary(middle_school_temp)
withfeatures_satisfied_09_15 <-
  withfeatures_satisfied_09_15 %>%
  select(-`feeder_dbn(june)`) %>%
  left_join(middle_school_temp, by="feeder_dbn")
summary(withfeatures_satisfied_09_15)
rm(middle_school_temp)
```

#### Treat specialized codes

Here, using `final_disposition_dbn` and `final_disposition`, we are able to create a table with school dbns and program codes for 9 specialized HS

```{r specialized-codes}

# Across ALL years, LaGuardia DOES have program codes, at least in final_disposition
withfeatures_satisfied_09_15 %>% filter(final_disposition_dbn == "03M485") %>% count(year, final_disposition) # == "03M485")

# check for program codes in the 9 Specialized High Schools. They do exist, and do exist through all the years.
withfeatures_satisfied_09_15 %>% 
  filter(substring(final_disposition_dbn, 3) %in% c("X445", "K449", "K430", "R605", "M475", "M692", "X696", "Q687", "M485")) %>%
  count(year, final_disposition)

# that's why now we create a table for specialized program codes
library(readr)
names2016 <- read_csv(url('https://data.cityofnewyork.us/api/views/7crd-d9xh/rows.csv?accessType=DOWNLOAD'))
specialized_code_school_mapping <-
  withfeatures_satisfied_09_15 %>% 
  filter(substring(final_disposition_dbn, 3) %in% c("X445", "K449", "K430", "R605", "M475", "M692", "X696", "Q687", "M485")) %>%
  group_by(final_disposition_dbn) %>%
  select("specialized_dbn" = final_disposition_dbn, 
         "specialized_code" = final_disposition) %>%
  distinct(specialized_code, .keep_all = TRUE) %>%
  arrange(specialized_dbn) %>%
  left_join(names2016, by=c("specialized_dbn" = "dbn")) %>%
  select(school_name, specialized_dbn, specialized_code)
save(specialized_code_school_mapping, file ="/data/nycdoe/clean_data/specialized_code_school_mapping.Rdata")
rm(names2016)

# apparently, students CAN NOT list any of the specialized codes above in one of their choices
withfeatures_satisfied_09_15 %>% 
  filter(r1programcode1 %in% specialized_code_school_mapping$specialized_code | 
           r1programcode2 %in% specialized_code_school_mapping$specialized_code | 
           r1programcode3 %in% specialized_code_school_mapping$specialized_code)

# add specialized schools to the fuller table 
spec_codes <- 
  specialized_code_school_mapping %>%
  ungroup() %>%
  mutate(pgname = rep("Missing", times=nrow(specialized_code_school_mapping)),
         interest = rep("Missing", times=nrow(specialized_code_school_mapping)),
         method = rep("Missing", times=nrow(specialized_code_school_mapping))) %>%
  select("dbn" = specialized_dbn,
         "pgcode" = specialized_code,
         pgname, school_name, interest, method)
percode_schoolmapping_fuller <-
  rbind(percode_schoolmapping_fuller, spec_codes)
summary(percode_schoolmapping_fuller)
rm(spec_codes)
save(percode_schoolmapping_fuller, file = '/data/nycdoe/clean_data/schoolcodemapping_fuller.Rdata')
```

#### Map program codes in top1 choice to schools

```{r map-programs-to-schools}

# what frac not having top1 choice (NA): 0.567% (2754)
withfeatures_satisfied_09_15 %>% count(is.na(r1programcode1))

# among these 2754:
# 519 applied to lag
# 2000 took shsat test
# 235 both applied to lag AND took shsat test
withfeatures_satisfied_09_15 %>%
  filter(is.na(r1programcode1)) %>%
  mutate(lag_applied = as.numeric(lag_arts_applied==1 | lag_dance_applied==1 | lag_instr_applied==1 | lag_vocal_applied==1 | lag_theatre_applied==1 | lag_drama_applied==1)) %>%
  count(shsat_tested, lag_applied)

# FOR NOW, we will drop these students, because we need to r1r1 include their high school features in the model.
withfeatures_fullr1r1_09_15 <-
  withfeatures_satisfied_09_15 %>% filter(!is.na(r1programcode1))
# BUT as for the original question "Do you get what you want?", we can still evaluate whether or not these are satisfied.
# One thing we can do is to create a proxy. For any students applying to Specialized programs and not listing top choice(s), we use the proxy to fill in instead.
# The proxy is identified by:
## (1) create a list of programs/schools that Specialized applicants tend to list as their top choices
## (2) plot the distribution of applications to these schools

# map programcodes to schools!
withfeatures_fullr1r1_09_15 <-
  withfeatures_fullr1r1_09_15 %>%
  #select(-(r1r1_pgname:r1r1_method)) %>%
  left_join(., percode_schoolmapping_fuller, by=c("r1programcode1" = "pgcode")) %>% #, suffix=c("", ".y")) %>%
  rename("r1r1_pgname" = pgname,
         "r1r1_dbn" = dbn,
         "r1r1_schoolname" = school_name,
         "r1r1_interest" = interest,
         "r1r1_method" = method)
```

#### High school's features 

```{r add-high-school-features}

# there are 152 `r1programcode1` that are NOT in `hs_features_new` - this is 10,986 (2.274%) out of all applications!
withfeatures_fullr1r1_09_15 %>%
  filter(!(r1r1_dbn %in% hs_features_new$dbn)) #%>% distinct(r1programcode1) 

# there are almost 300 HIGH schools in `hs_features_new` that do not exist in the full code-school map. Because we map `r1programcode1` to `r1r1_dbn` through the `percode_schoolmapping_fuller` table, this missing-out means there are choice programs that don't have a chance to be connected to school features.
hs_features_new %>%
  filter(!is.na(English_regents)) %>%
  filter(!(dbn %in% percode_schoolmapping_fuller$dbn))

# add HIGH SCHOOL features
high_school_temp <- hs_features_new
names(high_school_temp) <- paste("r1r1", names(high_school_temp), sep="_")
withfeatures_fullr1r1_09_15 <-
  withfeatures_fullr1r1_09_15 %>%
  #select(-(41:48)) %>%
  left_join(high_school_temp, by="r1r1_dbn")

# there are 10,986 rows with NA's in HS demographics (because r1programcode1 can't be mapped to any school) and 13,896 rows with NA's in regents.
summary(withfeatures_fullr1r1_09_15)
rm(high_school_temp)

# these extra 2,910 applications applied to 14 schools. In other words, these 14 schools do not have regents data.
withfeatures_fullr1r1_09_15 %>%
  filter(!is.na(r1r1_perc_poverty)) %>%
  filter(is.na(r1r1_English_regents)) %>%
  count(r1r1_dbn)

# drop 13,896 rows with NA's in high school's regents. This already encompasses the ~11k rows with NA's in demographics and attendance.
withfeatures_fullr1r1_09_15 <- 
  withfeatures_fullr1r1_09_15 %>%
  filter(!is.na(r1r1_English_regents))
```

### Start our regression!

```{r save-some-finalized-tables}

## SATISFIED, INCL. SPECIALIZED CASES #######
# has a column indicating whether or not an applicant is satisfied, i.e. get offered from the specialized program they apply to, OR get finalized to their top1 choie. 
# students in this table have to participate in at least ONE of these 3 activities: (1) applied to any LaGuardia program, (2) took the SHSAT test, and (3) listed top1 choice.
satisfied_laguardia_specialized_top1_09_15 <- satisfied_09_15
save(satisfied_laguardia_specialized_top1_09_15, 
     file = "/data/nycdoe/clean_data/satisfied_laguardia_specialized_top1_09_15.Rdata")


## STUDENTS WITH TOP3 CHOICES #######
# to be used for "school recommendation" networks
students_w_top3_choices_09_15 <- 
  withfeatures_satisfied_09_15 %>%
  select(year, student_id_scram, r1programcode1, r1programcode2, r1programcode3, shsat_tested, shsat_offer, any_lag_offer, starts_with("lag"), lag_offered, was_satisfied, grade_level, borough, feeder_borough, sex, ethnicity, ell, swd, poverty, GPA, quantiled_GPA)
save(students_w_top3_choices_09_15, file ="/data/nycdoe/clean_data/students_w_top3_choices_09_15.Rdata")


## STUDENTS WITH FEATURES #######
# ensure applicants in 8th grade and feeder_dbn matches 8th-gr dbn in June Biog
save(withfeatures_satisfied_09_15, file ="/data/nycdoe/clean_data/withfeatures_satisfied_09_15.Rdata")


## NO MISSING TOP1 CHOICE #######
# in addition to Student-With-Features above, ensure applicants filled in their top1 choie, no matter if they applied to any Specialized program or not. This means those who ONLY applied to Specialized prorams and NOT list choices are disregarded.
save(withfeatures_fullr1r1_09_15, file ="/data/nycdoe/clean_data/withfeatures_fullr1r1_09_15.Rdata")
```

```{r create-model, message=FALSE}

library(glmnet)
library(scales)
library(ROCR)
library(Matrix)
library(modelr)
theme_set(theme_bw())
options(repr.plot.width=4, repr.plot.height=3)

model_data <- withfeatures_fullr1r1_09_15

# because there are 2325 NA's in graduation rate of high schools, we will NOT include this feature in the model
model_data %>% count(is.na(r1r1_perc_grad))

frmula <- as.formula(was_satisfied ~ borough + sex + ethnicity + ell + swd + poverty + quantiled_GPA + 
                                     feeder_attendance + feeder_total_stud + feeder_perc_poverty + 
                                     feeder_perc_White + feeder_perc_Black + feeder_perc_Hispanic + 
                                     feeder_perc_MultiRacial + feeder_perc_NativeAmer + feeder_perc_Asian +
                                     feeder_borough + r1programtype1 + r1r1_interest + r1r1_English_regents + 
                                     r1r1_Math_regents + r1r1_History_regents + r1r1_Science_regents +
                                     r1r1_attendance + r1r1_total_stud + r1r1_perc_poverty + 
                                     r1r1_perc_White + r1r1_perc_Black + r1r1_perc_Hispanic + 
                                     r1r1_perc_MultiRacial + r1r1_perc_NativeAmer + r1r1_perc_Asian)
mooo <- model.matrix(object = frmula, 
                     data = model_data)

# this take almost 5 minutes to run:
logit_satisfied <- glmnet(x = mooo,
                          y = model_data$was_satisfied,
                          family = "binomial", lambda = 0)   #lambda = 0 means no regularization
plot(logit_satisfied)

### INVESTIGATE COEFFICIENTS #####
coef(logit_satisfied)

library(broom)
tidy(logit_satisfied)

# features that matter the most
tidy(logit_satisfied) %>% 
  arrange(estimate) %>%
  tail(10) %>%
  ggplot(aes(y=term, x=estimate)) +
  geom_point() +
  labs(title = "Features that matter the most")# + scale_x_log10()
# features that matter the least
tidy(logit_satisfied) %>% 
  arrange(estimate) %>%
  head(10) %>%
  ggplot(aes(y=term, x=estimate)) +
  geom_point() +
  labs(title = "Features that matter the least")
```

```{r model-evaluation, message=FALSE}
# baseline-accuracy means that if you always predict ppl to be satisfied, you are right 49% times.
(baseline_accuracy <- mean(model_data$was_satisfied))
  
df <- data.frame(actual = model_data$was_satisfied,
                 predicted = predict(object = logit_satisfied, newx = mooo, type = "response"))  #, s = c(0.015, 0.01, 0.005)
summary(df)
df <-
  df %>%
  rename("predicted" = s0)
df <-
  df %>%
  mutate(adjusted_pred = ifelse(predicted > 0.5, yes=1, no=0))
summary(df)

# plot histogram of predicted probabilities to evaluate overconfidence
ggplot(df, aes(x = predicted)) +
  geom_histogram(binwidth = 0.01) +
  xlab('Predicted probability of satisfied') +
  ylab('Number of examples')

### CALIBRATION PLOT ######
df %>% 
  group_by(predicted=round(predicted*10)/10) %>%
  summarize(num=n(), actual=mean(actual == 1)) %>%
  ggplot(data=., aes(x=predicted, y=actual, size=num)) +
  geom_point() +
  geom_abline(linetype=2) +
  scale_x_continuous(labels=percent, lim=c(0,1)) +
  scale_y_continuous(labels=percent, lim=c(0,1)) +
  labs(x = 'Predicted probability of satisfied', 
       y = 'Percent that are actually satisfied',
       size = "Number of applications",
       title = "Calibration Plot")

### ROC curve and AUC #######
# create a ROCR object
pred <- prediction(df$predicted, df$actual)
# plot ROC curve
perf <- performance(pred, measure='tpr', x.measure='fpr')
plot(perf, main = "ROC curve for k=1 classifier")
# more fancy ROC
roc <- data.frame(fpr=unlist(perf@x.values), tpr=unlist(perf@y.values))
roc %>%
  ggplot(data=., aes(x=fpr, y=tpr)) + 
  geom_line() +
  geom_abline(linetype=2) +
  scale_x_continuous(labels=percent, lim=c(0,1)) +
  scale_y_continuous(labels=percent, lim=c(0,1)) +
  xlab("Probability of a false optimism") +
  ylab("Probability of identifying true satisfaction") +
  labs(title = "ROC curve for k=1 classifier") +
  theme(legend.position=c(0.7,0.2), legend.title=element_blank())
# y.values is the area under the curve!
performance(pred, 'auc')

actual_negatives <- nrow(df) - sum(df$actual)
actual_positives <- sum(df$actual)

false_negatives <- nrow(df[which(df$adjusted_pred == 0 & df$actual == 1), ])
false_positives <- nrow(df[which(df$adjusted_pred == 1 & df$actual == 0), ])

true_negatives <- nrow(df[which(df$adjusted_pred == 0 & df$actual == 0), ])
true_positives <- nrow(df[which(df$adjusted_pred == 1 & df$actual == 1), ])

# ACCURACY: fraction of correct classifications
(accuracy <- mean(df$adjusted_pred == df$actual))

# RECALL, SENSITIVITY: the model identifies 40% of actual-satisfied as satisfied
(recall <- true_positives / (true_positives + false_negatives)) 

# PRECISION: fraction of positive predictions that are actually true
(precision <- true_positives / (true_positives + false_positives))

# FALSE POSITIVE RATE: 
(false_positive_rate <- false_positives / (false_positives + true_negatives))
confMatrix <-
  matrix(c(true_positives, false_negatives,
           false_positives, true_negatives),
         nrow = 2,
         ncol = 2,
         byrow = TRUE,
         dimnames = list(c("actual=T", "actual=F"),
                         c("predicted=T", "predicted=F")))
confMatrix

mtrics <- data.frame(baseline_acc = baseline_accuracy,
                   accuracy = accuracy,
                   recall = recall,
                   precision = precision,
                   false_positive_rate = false_positive_rate)
mtrics
```
