---
title: "Where Do students go to school?? RO"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(ggplot2)
library(readr)
library(tidyverse)
#library(tigris)
library(dplyr)
library(leaflet)
library(sp)
library(ggmap)
library(maptools)
library(broom)
library(httr)
library(rgdal)
library(reshape2)
library(leaflet.extras)

files <- Sys.glob('/data/nycdoe/Zoned DBNs/*.csv')

read_csv_with_year <- function(filename) {
  year <- as.numeric(gsub('-.*$', '', basename(filename)))
  df <- read_csv(filename, col_types = cols(res_zip_cde = col_character(), census_block  = col_character(), census_tract = col_character()))
  df$year <- year
  df
}

zoneData <- map_df(files, read_csv_with_year)

elementary_zoneData <- zoneData %>% filter(!is.na(zoned_elm_dbn))
#records dropped: 152198
nrow(zoneData) - nrow(elementary_zoneData)
middle_zoneData <- zoneData %>% filter(!is.na(zoned_mid_dbn))
nrow(zoneData) - nrow(middle_zoneData)
#records dropped: 154924
hs_zoneData <- zoneData %>% filter(!is.na(zoned_hs_dbn))
nrow(zoneData) - nrow(hs_zoneData)
#records dropped 206430

##################################

files <- Sys.glob('/data/nycdoe/June Biog/*.csv')

read_csv_with_year <- function(filename) {
  year <- as.numeric(gsub('-.*$', '', basename(filename)))
  df <- read_csv(filename, col_types = cols(grade_level=col_character(), grade_code=col_character(), admit_code=col_character(), adcode=col_character(),fall_days_present = col_character(), spring_days_present = col_character()))
  df$year <- year
  df
}

bios <- map_df(files, read_csv_with_year)

```


<!-- #Join the bio and and zone data for elementary school. This is only from 2015. -->
<!-- ```{R echo = FALSE} -->
<!-- #elem students merge with bio -->

<!-- elem_year <- elementary_zoneData %>% filter(year == 2015) -->

<!-- dbnBioData <- bios %>% -->
<!--   select(student_id_scram, dbn, grade_level, year) %>% -->
<!--   filter(year == 2015) -->

<!-- dbnBioData <-  rename(dbnBioData, BioDbn = dbn) -->

<!-- elementary_zone_bio_join <- inner_join(dbnBioData, elem_year, by = "student_id_scram" ) -->

<!-- elementary_zone_bio <- elementary_zone_bio_join -->
<!-- ``` -->

# <span style="color:blue"> Tidy bio data </span>
###Map grades to school level to detirmine elementary, middle and high school
We assumed:

* k-5 is elementary school
* 6-8 is middle school
* 9-12 is high school
Calculations were made with this in mind. We are aware that some schools are only k-2, so students will be zoned differently for grade 3, but we are interested in the zone of kindergarten/lowest grade a students entered the system.

```{R} 
grades <- c(0, 01, 02, 03, 04, 05, 06, 07, 08, 09, 10, 11, 12)
schools <-c("em", "em", "em" , "em" , "em", "em", "mid" , "mid", "mid", "hs", "hs", "hs",  "hs")
map_grade <- data.frame(grades, schools)
```

###'regrade' the grade_level so that all grades have a numeric value. 
K is 0, all other grades before K are dropped because students are not zoned for school until Kindergarten.
```{R}
all_bios <- bios

#bios_noNA <- all_bios%>% 
 # filter(!is.na(student_id_scram)) #only one student id is na

bios_dbn <- all_bios %>% 
  select(student_id_scram, dbn, year, grade_level)

bios_dbn <- bios_dbn %>% mutate(grade_level = ifelse(grade_level == "0K", "00", grade_level))
bios_dbn <- bios_dbn %>% mutate(grade_level = as.integer(grade_level)) %>% filter(!is.na(grade_level))

#join data so you see students
bios_dbn <- bios_dbn %>% full_join(map_grade, by = c("grade_level" = "grades"))
################################ MOST IMPORTANT ###############################################
test<- bios_dbn %>% group_by(student_id_scram, schools, year, dbn) %>% summarise() 

#only keep entry to each school level for each student
bio_distinct_schools <- test %>% group_by(student_id_scram, schools) %>% filter(year == min(year))
####################################################################################################

#view
#bios_dbn %>% group_by(student_id_scram, dbn, schools) %>% summarise()%>% arrange(student_id_scram) %>% View()

smaller_data <- bios_dbn %>% group_by(student_id_scram, dbn, schools) %>% summarise()

unique_data <- smaller_data %>% group_by(student_id_scram, schools) %>% 
  distinct(student_id_scram, schools, .keep_all = TRUE) 

#it works!
# unique_data %>% 
#   group_by(student_id_scram) %>%
#   summarise(count =n())%>%
#   group_by(count) %>%
#   summarise(countx = n())

#spread this now to get distance from zone to school. IMPORTANT: We only took one mid, high, em for each student. 
#Some students changed schools, and we would not see this from this table.
#################################################################
bios_spread <- unique_data %>% spread(schools, dbn)
##################################################################

#display bio_distinct_schools

bio_distinct_schools

```
#Tidy zone data
```{R}
#############################
#ZONE DATA DOES NOT INCLUDE ANYTHING FROM 2005 or 2006
#############################

#sanity checks are commented

# zoneData %>% select(student_id_scram, zoned_elm_dbn, zoned_mid_dbn, zoned_hs_dbn, year) %>%
#   group_by(student_id_scram) %>% summarise(count = n()) %>% group_by(count) %>% summarise(countx = n()) 

# zoneData %>% select(student_id_scram, zoned_elm_dbn, zoned_mid_dbn, zoned_hs_dbn, year) %>%
#   group_by(student_id_scram) %>% summarise(count = n()) %>% 
#   filter(count == 9) %>% head(1)

#zoneData %>% filter(student_id_scram == 12507)

distinct_zones <- zoneData %>% select(1:8) %>%  #used to be callued df_distinct
  group_by(student_id_scram, zoned_elm_dbn, zoned_mid_dbn, zoned_hs_dbn, year) %>% 
  distinct(student_id_scram, schools, .keep_all = TRUE) 

#df_distinct %>% group_by(year) %>% summarise()

distinct_zones
```
#Join zone and bio data on DBN
We joined the data on DBNs from the zoned data and the DBN's from the bio data on each elementary, middle and high school.
It is important to note that <span style = "color:red"><b>zone data only begins from 2007</b></span>. 
For students who began elementary, middle or high school before 2007, we do NOT have the zoned dbn. Therefore, naturally these students are not included on maps of attended school vs zoned school. For all other counts, these students are included.

<span style="color:red">Data Quirks:</span> Students are zoned to an elementary, middle and high school every time he/she is zoned. We had to detirmine the year a student entered elementary school, for example, and then find that year for that student in the zoned data to detirmine the zone that student was mapped to the year he/she entered elementary school. Zones change quite frequently and this granular data was important for the numbers later on.

```{R}
# reformat df to wide format
bio_distinct_schools <- bio_distinct_schools %>% spread(schools, dbn)

#join the data frames. All records that do not have a match are dropped. This ensures that the map will only include data that exists
join_zone_bio <- inner_join(bio_distinct_schools, distinct_zones, by = c("year", "student_id_scram"))

join_zone_bio
```


<!-- only 2015 -->
<!-- ```{R echo = FALSE} -->
<!-- #summarise the grade levels -->
<!-- elementary_zone_bio <- elementary_zone_bio %>% mutate(grade_level = ifelse(grade_level == "0K", "00", grade_level)) -->

<!-- # convert grades to integers where 0 = Kindergarten -->
<!-- elementary_zone_bio <- elementary_zone_bio %>% mutate(grade_level = as.integer(grade_level)) %>% -->
<!--   filter(!is.na(grade_level))  -->

<!-- elementary_zone_bio <- elementary_zone_bio %>% filter(grade_level<6) %>% select(year.x, BioDbn, zoned_elm_dbn, student_id_scram, grade_level) -->

<!-- elem_zone_tidy <- elementary_zone_bio %>% select(student_id_scram, zoned_elm_dbn, BioDbn, grade_level) -->

<!-- ``` -->
<!-- Match -->
<!-- ```{r echo = FALSE} -->
<!-- # match on zoned school and bio -->
<!-- elem_zone_tidy %>% mutate(match_dbn = ifelse(zoned_elm_dbn == BioDbn, 1, 0)) %>%  -->
<!--   group_by(match_dbn) %>% summarise(count = n()) -->

<!-- # group by grade and get counts of matches by grade -->
<!-- elem_zone_tidy %>% mutate(match_dbn = ifelse(zoned_elm_dbn == BioDbn, 1, 0)) %>%  -->
<!--   group_by(grade_level, match_dbn) %>% summarise(count = n()) -->


<!-- match_df <- elem_zone_tidy %>% mutate(match_dbn = ifelse(zoned_elm_dbn == BioDbn, 1, 0)) -->

<!-- ``` -->

#Match zoned and attended schools
* 0 has em dbn, but does not match
* 1 has em dbn and matches
* -1 does NOT have em dbn
```{r}
join_zone_bio_matches <- join_zone_bio

join_zone_bio_matches <- join_zone_bio_matches%>% 
  mutate(matchEM = ifelse(!is.na(zoned_elm_dbn), ifelse(zoned_elm_dbn == em, 1, 0), -1))

join_zone_bio_matches <- join_zone_bio_matches%>% 
  mutate(matchMID = ifelse(!is.na(zoned_mid_dbn), ifelse(zoned_mid_dbn == mid, 1, 0), -1))

join_zone_bio_matches <- join_zone_bio_matches%>% 
  mutate(matchHS = ifelse(!is.na(zoned_hs_dbn), ifelse(zoned_hs_dbn == hs, 1, 0), -1))
```
#HS school data

add a collumn for boro
* note that some DBNs are not in <b>district-boro-n<b> format. 
```{r}
# some dbns don't have a 0 boro and the dbn is in not in standard dbn format
# 69,974 have a zoned boro of zero, 168,331 have a standard zoned boro

hs_data <- join_zone_bio_matches %>%filter(!is.na(hs)) %>% mutate(zoned_boro = substr(zoned_hs_dbn, 3,3))
hs_data

#%>% 

  #group_by(zoned_boro) %>% 
  #  summarise(count = n())
  
```

## HS Students
```{r}
# number of students zoned to a school (includes all years of data), we don't know how many seats each school has
hs_data %>% group_by(zoned_hs_dbn, year) %>% summarise(num_zoned_to_dbn = n()) 

```

```{r}
# number of students who went to a specific elementary school
local_zones_data <- full_join( 
  (hs_data %>% group_by(zoned_hs_dbn) %>% summarise(zoned_to_dbn = n())), 
  (hs_data %>% group_by(hs) %>% summarise(went_to_dbn = n())),
           by = c("zoned_hs_dbn" = "hs") )

local_zones_data <- full_join( 
  local_zones_data, 
  (join_zone_bio_matches %>% filter(matchHS == 1) %>% 
     group_by(zoned_hs_dbn) %>% 
   summarise(attending_zoned_dbn = n())),   by = "zoned_hs_dbn")

local_zones_data <- local_zones_data %>% mutate(perc_local = (attending_zoned_dbn/went_to_dbn)*100)
local_zones_data %>% mutate(boro = substr(zoned_hs_dbn, 3, 3)) %>% filter(!is.na(attending_zoned_dbn)) %>% group_by(boro) %>% summarize(sum_attended_zoned=sum(attending_zoned_dbn), sum_zoned= sum(zoned_to_dbn)) %>%
  mutate(perc_boro_att_zoned = (sum_attended_zoned/sum_zoned)*100)


local_zones_data

```

```{r}
#NA Schools

# Closed schools:
# 17K440 - Prospect Heights HS (2008)
# 11X425 - Evander Childs HS (2006)
# 17k470 - George Wingate HS (2006)
# 19K435 - THOMAS JEFFERSON HS

#67 schools had no students go to that school
#schools in district 71 do not seem to appear in open data set
local_zones_data %>% filter(is.na(went_to_dbn))

#dropping these 67 schools
local_zones_data %>% filter(!is.na(went_to_dbn))

#17 schools that only have one student that went to that school
#3,508 students zoned to 18k515 but it has closed - South Shore High School
local_zones_data %>% filter(went_to_dbn == 1)

#dropping these schools
local_zones_data %>% filter(!went_to_dbn == 1)

#742 schools where no one was zoned but students went there
local_zones_data %>% filter(is.na(zoned_to_dbn))

#final df: 742 schools out of 768 schools = no one was zoned but students went there
local_zones_data %>% filter(!went_to_dbn == 1, !is.na(went_to_dbn))

nrow(local_zones_data)

```







#MS school data

add a collumn for boro
* note that some DBNs are not in <b>district-boro-n<b> format. 
```{r}
# some dbns don't have a 0 boro and the dbn is in not in standard dbn format
# 69,974 have a zoned boro of zero, 168,331 have a standard zoned boro

mid_data <- join_zone_bio_matches %>%filter(!is.na(mid)) %>% mutate(zoned_boro = substr(zoned_mid_dbn, 3,3))
mid_data

#%>% 

  #group_by(zoned_boro) %>% 
  #  summarise(count = n())
  
```

## MS Students
```{r}
# number of students zoned to a school (includes all years of data), we don't know how many seats each school has
mid_data %>% group_by(zoned_mid_dbn, year) %>% summarise(num_zoned_to_dbn = n()) 

```

```{r}
# number of students who went to a specific elementary school
mid_local_zones_data <- full_join( 
  (mid_data %>% group_by(zoned_mid_dbn) %>% summarise(zoned_to_dbn = n())), 
  (mid_data %>% group_by(mid) %>% summarise(went_to_dbn = n())),
           by = c("zoned_mid_dbn" = "mid") )

mid_local_zones_data <- full_join( 
  mid_local_zones_data, 
  (join_zone_bio_matches %>% filter(matchMID == 1) %>% 
     group_by(zoned_mid_dbn) %>% 
   summarise(attending_zoned_dbn = n())),   by = "zoned_mid_dbn")

mid_local_zones_data <- mid_local_zones_data %>% mutate(perc_local = (attending_zoned_dbn/went_to_dbn)*100)

mid_local_zones_data %>% mutate(boro = substr(zoned_mid_dbn, 3, 3)) %>% filter(!is.na(attending_zoned_dbn)) %>% group_by(boro) %>% summarize(sum_attended_zoned=sum(attending_zoned_dbn), sum_zoned= sum(zoned_to_dbn)) %>%
  mutate(perc_boro_att_zoned = (sum_attended_zoned/sum_zoned)*100)


mid_local_zones_data

```

MS
```{r}
#NA Schools

#1,050 schools
nrow(mid_local_zones_data)

#175 schools had no students go to that school
#of those, 111 are not zone schools, others seem to have closed.
mid_local_zones_data %>% filter(is.na(went_to_dbn))

#15 schools that only have one student that went to that school
mid_local_zones_data %>% filter(went_to_dbn == 1)

#665 schools where no one was zoned but students went there
mid_local_zones_data %>% filter(is.na(zoned_to_dbn))

#final df: 665 schools out of 860 schools = no one was zoned but students went there
mid_local_zones_data %>% filter(!went_to_dbn == 1, !is.na(went_to_dbn))

```





#EM school data

```{r}
# some dbns don't have a 0 boro and the dbn is in not in standard dbn format
# 69,974 have a zoned boro of zero, 168,331 have a standard zoned boro

em_data <- join_zone_bio_matches %>%filter(!is.na(em)) %>% mutate(zoned_boro = substr(zoned_elm_dbn, 3,3))
em_data

#%>% 

  #group_by(zoned_boro) %>% 
  #  summarise(count = n())
  
```

## MS Students
```{r}
# number of students zoned to a school (includes all years of data), we don't know how many seats each school has
em_data %>% group_by(zoned_elm_dbn, year) %>% summarise(num_zoned_to_dbn = n()) 

```

```{r}
# number of students who went to a specific elementary school
em_local_zones_data <- full_join( 
  (em_data %>% group_by(zoned_elm_dbn) %>% summarise(zoned_to_dbn = n())), 
  (em_data %>% group_by(em) %>% summarise(went_to_dbn = n())),
           by = c("zoned_elm_dbn" = "em") )

em_local_zones_data <- full_join( 
  em_local_zones_data, 
  (join_zone_bio_matches %>% filter(matchEM == 1) %>% 
     group_by(zoned_elm_dbn) %>% 
   summarise(attending_zoned_dbn = n())),   by = "zoned_elm_dbn")

em_local_zones_data <- em_local_zones_data %>% mutate(perc_local = (attending_zoned_dbn/went_to_dbn)*100)

em_local_zones_data %>% mutate(boro = substr(zoned_elm_dbn, 3, 3)) %>% filter(!is.na(attending_zoned_dbn)) %>% 
  group_by(boro) %>% 
  summarize(sum_attended_zoned=sum(attending_zoned_dbn), sum_zoned= sum(zoned_to_dbn)) %>%
  mutate(perc_boro_att_zoned = (sum_attended_zoned/sum_zoned)*100)


em_local_zones_data

```



```{r}
#NA Schools

#1,313 schools
nrow(em_local_zones_data)

#200 schools had no students go to that school
em_local_zones_data %>% filter(is.na(went_to_dbn))

#11 schools that only have one student that went to that school
em_local_zones_data %>% filter(went_to_dbn == 1)

#425 schools where no one was zoned but students went there
em_local_zones_data %>% filter(is.na(zoned_to_dbn))

#final df:
em_local_zones_data %>% filter(!went_to_dbn == 1, !is.na(went_to_dbn))

```




