---
title: "Final Plots"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(dplyr)
library(tidyr)
library(ggplot2)
library(reshape2)
library(boot)
library(glmnet)
library(tidyverse)
library(scales)
library(ROCR)
library(directlabels)

theme_set(theme_bw())
```

## Student Drop Outs Throughout High School
```{r}
load('/data/nycdoe/clean_data/hs_student_dropouts_data.Rdata')

student_drop_outs_throughout_hs <- final_df_2 %>%
  filter(year<=2012) %>%
  ggplot(aes(x=grade, y=num_students, color=as.factor(year))) +
  geom_point() +
  geom_line() +
  scale_y_continuous(labels = percent) +
  scale_x_continuous(breaks=seq(9,13,1), labels=c('9'='9', '10'='10', '11'='11', '12'='12', '13'='Graduated'), expand=c(0, 0.75)) +
  geom_dl(aes(label = as.factor(year)), method = list(dl.trans(x = x + .3), "last.points", cex = 0.8)) +
  labs(x="Grade Level", y="Percent of Incoming Freshmen", title="Student Drop-Outs Throughout High School", color="Cohort of")

student_drop_outs_throughout_hs
```

## Student Drop Outs Throughout High School - Faceted by Ethnicity
```{r}
load('/data/nycdoe/clean_data/hs_student_dropouts_eth_data.Rdata')

final_df_2$ethnicity <- as.factor(final_df_2$ethnicity)
levels(final_df_2$ethnicity) <- c("No Ethnicity Listed", "Asian", "Black", "Hispanic", "Multi-Racial", "Native American", "White")

student_drop_outs_throughout_hs_ethnicity <- final_df_2 %>%
  filter(year<=2012) %>%
  ggplot(aes(x=grade, y=num_students, color=as.factor(year))) +
  geom_point() +
  geom_line() +
  scale_y_continuous(labels = percent) +
  scale_x_continuous(breaks=seq(9,13,1), labels=c('9'='9', '10'='10', '11'='11', '12'='12', '13'='Grad.'), expand=c(0, 0.75)) +
  geom_dl(aes(label = as.factor(year)), method = list(dl.trans(x = x + .3), "last.points", cex = 0.8)) +
  labs(x="Grade Level", y="Percent of Incoming Freshmen", title="Student Drop-Outs Throughout High School", color="Cohort of") +
  facet_wrap(~ethnicity)

student_drop_outs_throughout_hs_ethnicity
```

## Student Drop Outs Throughout High School - Faceted by Gender
```{r}
load('/data/nycdoe/clean_data/hs_student_dropouts_gender_data.Rdata')

student_drop_outs_throughout_hs_gender <- final_df_2 %>%
  filter(year<=2012) %>%
  ggplot(aes(x=grade, y=num_students, color=as.factor(year))) +
  geom_point() +
  geom_line() +
  scale_y_continuous(labels = percent) +
  scale_x_continuous(breaks=seq(9,13,1), labels=c('9'='9', '10'='10', '11'='11', '12'='12', '13'='Graduated'), expand=c(0, 0.75)) +
  geom_dl(aes(label = as.factor(year)), method = list(dl.trans(x = x + .3), "last.points", cex = 0.8)) +
  labs(x="Grade Level", y="Percent of Incoming Freshmen", title="Student Drop-Outs Throughout High School", color="Cohort of") +
  facet_wrap(~sex)

student_drop_outs_throughout_hs_gender
```

## Average Grade-Wise Continuation Rates Across 2005-2014
```{r}
load('/data/nycdoe/clean_data/average_gradewise_continuation_rates.Rdata')

avg_grade_wise_continuation_rates <- continuation_rates %>%
  group_by(grade_level) %>%
  summarise(avg_continuation=mean(continuation_rate), se=sd(continuation_rate)/n()) %>%
  ggplot(aes(x=grade_level, y=avg_continuation)) +
  scale_x_continuous(breaks=seq(-1, 12, 1), labels=c('-1'='Pre-K', '0'='K', '1'='1', '2'='2', '3'='3', '4'='4', '5'='5', '6'='6', '7'='7', '8'='8', '9'='9', '10'='10', '11'='11', '12'='12'), expand=c(0, 0.75)) +
  scale_y_continuous(labels = percent) +
  geom_point() +
  geom_line() +
  labs(x="Grade Level", y="Average Percent Continuation", title="Year-to-Year Continuation") +
  geom_pointrange(aes(ymin = avg_continuation-se, ymax = avg_continuation+se))

avg_grade_wise_continuation_rates
```

## Average Grade-Wise Continuation Rates Split by Year
```{r}
avg_grade_wise_continuation_rates_by_year <- continuation_rates %>%
  filter(year!=2015) %>%
  mutate(grade_level=as.character(grade_level)) %>%
  mutate(grade_level=ifelse(grade_level=='-1','Pre-K',
                            ifelse(grade_level=='0', 'K',
                                   grade_level))) %>%
  ggplot(aes(x=year, y=continuation_rate, color=factor(grade_level, levels=c('Pre-K', 'K', '1', '2', '3', '4', '5', '6', '7', '8', '9', '10', '11', '12')))) +
  geom_point() +
  geom_line() +
  labs(x="Academic Year", y="Percent Continuation", title="Trends in Continuation Rates by Grade Level", color="Grade Level") +
  scale_y_continuous(labels = percent) +
  scale_x_continuous(breaks=seq(2005, 2014, 1), expand=c(0, 1)) +
  geom_dl(aes(label = as.factor(grade_level)), method = list(dl.trans(x = x + .3), "last.points", cex = 0.75))

avg_grade_wise_continuation_rates_by_year
```

# Model: did_not_continue ~ grade_level + sex + ethnicity + ell + poverty + perc_attendance + quantiled_GPA + school_avg_attendance + school_avg_grad + school_avg_English + school_avg_History + school_avg_Language + school_avg_Math + school_avg_Science

## Coefficients
> coef(model_5)
107 x 1 sparse Matrix of class "dgCMatrix"
                                              s0
(Intercept)                         -0.110625063
(Intercept)                          .          
grade_level1                         0.418244557
grade_level-1                       -0.074831362
grade_level10                        2.656616302
grade_level11                        1.874840792
grade_level12                        1.160446307
grade_level2                         0.319546452
grade_level3                         0.139231775
grade_level4                        -0.047714778
grade_level5                         0.220681755
grade_level6                         0.308763573
grade_level7                         0.189668389
grade_level8                         0.618494254
grade_level9                         2.100186840
sexM                                 0.183916486
ethnicityAsian                      -0.273433214
ethnicityBlack                       0.060348604
ethnicityHispanic                   -0.104540124
ethnicityMulti-Racial                0.156259207
ethnicityNative American            -0.016677044
ethnicityWhite                      -0.148575450
ell                                  0.362096565
poverty                             -0.177393737
perc_attendance10                   -0.321894580
perc_attendance100                  -3.957311098
perc_attendance20                   -0.522810358
perc_attendance30                   -0.792461426
perc_attendance40                   -1.102650795
perc_attendance50                   -1.436796287
perc_attendance60                   -1.835128123
perc_attendance70                   -2.303149960
perc_attendance80                   -2.828415891
perc_attendance90                   -3.406070502
quantiled_GPA10                      0.084605308
quantiled_GPA2                      -0.293418907
quantiled_GPA3                      -0.333660314
quantiled_GPA4                      -0.345121123
quantiled_GPA5                      -0.301056779
quantiled_GPA6                      -0.311366801
quantiled_GPA7                      -0.211128917
quantiled_GPA8                      -0.160189884
quantiled_GPA9                      -0.035671002
quantiled_GPAMissing                 0.205950212
school_avg_attendance10             -0.191433759
school_avg_attendance100             0.162172608
school_avg_attendance20             -0.883179829
school_avg_attendance30              0.300304752
school_avg_attendance40              0.122643347
school_avg_attendance50              0.463188626
school_avg_attendance60              0.259482354
school_avg_attendance70             -0.034656320
school_avg_attendance80              0.014018812
school_avg_attendance90             -0.001099207
school_avg_attendanceNot Applicable  0.272905853
school_avg_grad10                   -0.720056084
school_avg_grad100                  -3.205442931
school_avg_grad20                   -0.817164379
school_avg_grad30                   -0.663692559
school_avg_grad40                   -1.080333303
school_avg_grad50                   -1.285676932
school_avg_grad60                   -1.665819093
school_avg_grad70                   -1.807981993
school_avg_grad80                   -2.002536153
school_avg_grad90                   -2.504625701
school_avg_gradNot Applicable       -0.617160124
school_avg_English20                 0.440004006
school_avg_English30                 0.321231745
school_avg_English40                 0.066901780
school_avg_English50                 0.438120672
school_avg_English60                 0.556773788
school_avg_English70                 0.636096990
school_avg_English80                 0.451221722
school_avg_English90                 0.197881136
school_avg_EnglishMissing            0.189212645
school_avg_EnglishNot Applicable     0.243227114
school_avg_History40                 0.157669797
school_avg_History50                 0.069993443
school_avg_History60                 0.008650314
school_avg_History70                 0.137211464
school_avg_History80                -0.013869884
school_avg_History90                -0.011203772
school_avg_HistoryMissing           -0.070460997
school_avg_HistoryNot Applicable     0.003919970
school_avg_Language100               0.157527439
school_avg_Language30                0.035323122
school_avg_Language60                0.983180750
school_avg_Language70               -0.112429084
school_avg_Language80                0.035415857
school_avg_Language90                0.155547613
school_avg_LanguageMissing           0.195185103
school_avg_LanguageNot Applicable    0.141311968
school_avg_Math50                    0.662214766
school_avg_Math60                    0.706552063
school_avg_Math70                    0.816062328
school_avg_Math80                    0.875305681
school_avg_Math90                    0.765416242
school_avg_MathMissing               0.830381998
school_avg_MathNot Applicable        0.782409904
school_avg_Science40                 0.521784959
school_avg_Science50                 0.586960045
school_avg_Science60                 0.488471262
school_avg_Science70                 0.383724621
school_avg_Science80                 0.265078384
school_avg_Science90                 0.211097957
school_avg_ScienceMissing            0.476142627
school_avg_ScienceNot Applicable     0.397416778

## Model Stats

Accuracy
```{r}
load('/data/nycdoe/clean_data/model_5_stats.Rdata')

accuracy <-
  model_5_stats %>%
  summarise(accuracy=sum(num_correct)/sum(total), baseline_accuracy=sum(cont)/sum(total))

accuracy
```

Precision
```{r}
precision <-
  model_5_stats %>%
  filter(p_did_not_continue>=0.5) %>%
  summarise(prec=sum(did_not_cont)/sum(total))

precision
```

True Positive Rate
```{r}
tpr <-
  model_5_stats %>%
  summarise(tpr=sum(did_not_cont*(p_did_not_continue>=0.5))/sum(did_not_cont))

tpr
```

False Positive Rate
```{r}
fpr <-
  model_5_stats %>%
  summarise(fpr=sum(cont*(p_did_not_continue>=0.5))/sum(cont))

fpr
```

AUC and ROC
```{r}
auc_data <-
  model_5_stats %>%
  select(cont, did_not_cont, p_did_not_continue) %>%
  gather('label', 'count', 1:2) %>%
  mutate(label=ifelse(label=='cont', 0, 1)) %>%
  mutate(actual_did_not_continue=label) %>%
  select(actual_did_not_continue, p_did_not_continue, count)

auc_data <- auc_data[rep(row.names(auc_data), auc_data$count), 1:2]

pred <- prediction(auc_data$p_did_not_continue, auc_data$actual_did_not_continue)

perf_nb <- performance(pred, measure='tpr', x.measure='fpr')

plot(perf_nb)

performance(pred, 'auc')
```

## Calibration Plot
```{r}
calibration_plot <- model_5_stats %>%
  group_by(predicted=round(p_did_not_continue*10)/10) %>%
  summarise(num=sum(total), actual=sum(did_not_cont)/sum(total)) %>%
  ggplot(aes(x=predicted, y=actual, size=num)) +
  geom_point() +
  geom_abline(linetype=2) +
  scale_x_continuous(labels=percent, lim=c(0,1)) +
  scale_y_continuous(labels=percent, lim=c(0,1)) +
  scale_size_continuous(labels=comma) +
  labs(title='Calibration plot for \'Automated Guidance Counselor\' Classification Model', x='Predicted probability that a student will not continue', y= 'Actual probability that a student will not continue', size='Number of students')

calibration_plot

ggsave(filename = '../figures/gc_model_calibration_plot.png', width=6)
```

## What Percent of Dropouts Are We Catching?
```{r}
b <- data.frame(c(1,1))
b[1,1] <-
  model_5_stats %>%
  summarise(sum(did_not_cont)/sum(total))

precision_plot <- model_5_stats %>%
  arrange(desc(p_did_not_continue)) %>%
  mutate(cum_total=cumsum(total), perc_cum_total=cum_total/sum(total), cum_did_not_cont=cumsum(did_not_cont), perc_cum_did_not_cont=cum_did_not_cont/sum(did_not_cont)) %>%
  ggplot(aes(x=perc_cum_total, y=perc_cum_did_not_cont)) +
  geom_line() +
  scale_x_continuous(labels=percent, breaks = seq(0, 1, 0.1)) +
  scale_y_continuous(labels=percent) +
  labs(x="Percent of Students Predicted To Drop Out", y="Percent of At-Risk Students Caught")

precision_plot

ggsave(filename = '../figures/gc_model_precision_plot.png', height=4, width=4)
```

## How Many Students We Intervened With Weren't Actually Going to Drop Out? 'Top K Curve'
```{r}
baseline <-
  model_5_stats %>%
  summarise(base=sum(did_not_cont)/sum(total))

recall_plot <- model_5_stats %>%
  arrange(desc(p_did_not_continue)) %>%
  mutate(cum_total=cumsum(total), perc_cum_total=cum_total/sum(total), cum_did_not_cont=cumsum(did_not_cont), p_dropout_in_top_k=cum_did_not_cont/cum_total) %>%
  ggplot(aes(x=perc_cum_total, y=p_dropout_in_top_k)) +
  geom_line() +
  scale_x_continuous(labels=percent, breaks = seq(0, 1, 0.1)) +
  scale_y_continuous(labels=percent, limits=c(0,1)) +
  labs(x="Top k% of Students Most Likely To Drop Out (According To Model)", y="Percent of Students in Top k% Who Were Actually At Risk") +
  geom_hline(yintercept = baseline[1,1], linetype='dashed')

recall_plot

ggsave(filename = '../figures/gc_model_recall_plot.png', height=5, width=4)
```

```{r}
baseline <-
  model_5_stats %>%
  summarise(base=sum(did_not_cont)/sum(total))

recall_plot_2 <- model_5_stats %>%
  arrange(desc(p_did_not_continue)) %>%
  mutate(cum_total=cumsum(total), perc_cum_total=cum_total, cum_did_not_cont=cumsum(did_not_cont), p_dropout_in_top_k=cum_did_not_cont/cum_total) %>%
  ggplot(aes(x=perc_cum_total, y=p_dropout_in_top_k)) +
  geom_line() +
  scale_x_continuous(labels=comma) +
  scale_y_continuous(labels=percent, limits=c(0,1)) +
  labs(x="Top k Students Predicted To Drop Out", y="Percent of Students in Top k Who Were Actually At Risk") +
  geom_hline(yintercept = baseline[1,1], linetype='dashed')

recall_plot_2

ggsave(filename = '../figures/gc_model_recall_plot.png', height=5, width=4)
```

```{r}
ratio_plot <- model_5_stats %>%
  arrange(desc(p_did_not_continue)) %>%
  mutate(cum_total=cumsum(total), perc_cum_total=cum_total/sum(total), cum_did_not_cont=cumsum(did_not_cont), p_dropout_in_top_k=cum_did_not_cont/cum_total, perc_cum_did_not_cont=cum_did_not_cont/sum(did_not_cont), ratio=(1-p_dropout_in_top_k)/perc_cum_did_not_cont) %>%
  ggplot(aes(x=perc_cum_total, y=ratio)) +
  geom_line() +
  scale_x_continuous(labels=percent) +
  scale_y_continuous(labels=percent, limits=c(0.5,1)) +
  labs(x="Top k% Students Predicted To Drop Out", y="Ratio of False Positives to True Positives")

ratio_plot
```
## Our Model's Predictions for a 'Typical' Student - Faceted by Ethnicity and Gender

Typical Student:
ell=0,
poverty=1,
perc_attendance=100,
quantiled_GPA='Missing',
school_avg_attendance=90,
school_avg_grad=60,
school_avg_English=60,
school_avg_History=60,
school_avg_Language=90,
school_avg_Math=60,
school_avg_Science=60

```{r}
load('/data/nycdoe/clean_data/model_5_stats_eth_gender.Rdata')

model_predictions_for_typical_student <- model_5_stats_on_artificial_test_with_features_4 %>%
  group_by(grade_level, ethnicity, sex) %>%
  summarise(predicted_avg_dropout=mean(p_did_not_continue), actual_avg_dropout=mean(actual_avg_dropout)) %>%
  ggplot(aes(x=grade_level, y=predicted_avg_dropout, color=ethnicity, group=ethnicity)) +
  geom_line() +
  geom_point(aes(x=grade_level, y=actual_avg_dropout)) +
  facet_wrap(~ sex)

model_predictions_for_typical_student
```
