---
title: "more features"
author: "Keri and David"
date: "7/24/2017"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, root.dir = '/data/nycdoe/')
options(tibble.width=Inf)
library(ggplot2)
library(readr)
library(tidyverse)
library(reshape2)
library(stringr)

#Reading in buncha relevant data frames
load('/data/nycdoe/clean_data/student_perf_feature.Rdata')

```
Read in June Biog Data for grabbing year student_id and grade level 
Then combining these two data frames together 
```{r}
read_june_with_year <- function(filename) {
  year <- as.numeric(gsub('-.*$', '', basename(filename)))
  df <- read_csv(filename, col_types = cols(student_id_scram = col_character(), grade_level = col_integer()))
  df$year <- year
  df <- select(df, year, student_id_scram, grade_level)
}
files <- Sys.glob('/data/nycdoe/June Biog/*.csv')
grade_level <- map_df(files, read_june_with_year)

stud_perf <- full_join(grade_level, student_performance, by= c("year","student_id_scram"))

save(stud_perf, file = '/data/nycdoe/clean_data/stud_perf_v2.Rdata')
```
Quantile GPA
```{r}
stud_perf_GPA <-
  stud_perf %>%
  select(year,student_id_scram,GPA) %>%
  filter(!is.na(GPA))
stud_perf_GPA_v2<-
  stud_perf_GPA %>%
  within(quantiled_GPA <- as.integer(cut(GPA, quantile(GPA, probs=0:10/10), include.lowest=TRUE)))

stud_perf_quantGPA <- full_join(stud_perf, stud_perf_GPA_v2, by=c("year","student_id_scram","GPA"))

stud_perf_quantGPA <-
  stud_perf_quantGPA %>% mutate(quantiled_GPA = ifelse(is.na(quantiled_GPA), "Missing", quantiled_GPA))

save(stud_perf_quantGPA, file='/data/nycdoe/clean_data/stud_perf_quantiledGPA.Rdata')
#add NAs later
```
Quantile Regents 
```{r}
#load in regents data
load('/data/nycdoe/clean_data/student_regents_quantile.Rdata')
stud_perf_quantiled <- full_join(stud_perf_quantGPA, student_regents_quantile, by=c("year","student_id_scram"))
#combine regents quantile to full data set 
```
How many people test at a grade level different than their own
Only applies to: ELA, Math, Sci, and Soc Exams
```{r}
#Social Studies
graph1 <- stud_perf_quantGPA %>%
  select(year, student_id_scram,grade_level,soc_grade_level,soc_perf_level) %>%
  group_by(grade_level, soc_grade_level) %>%
  count() %>% 
  filter(!is.na(soc_grade_level)) %>% View
  ggplot(aes(x=soc_grade_level, y = n)) +
  geom_point() +
  facet_wrap(~grade_level) +
  labs(title = "Social Studies faceted by student grade level", x = "Exam Level", y = "Number of People who took the exam")
plot(graph1)

#Science
graph2 <- stud_perf_quantGPA %>%
  select(year, student_id_scram,grade_level,sci_grade_level,sci_perf_level) %>%
  group_by(grade_level, sci_grade_level) %>%
  count() %>% 
  filter(!is.na(sci_grade_level)) %>%
  ggplot(aes(x=sci_grade_level, y = n)) +
  geom_point() +
  facet_wrap(~grade_level)  +
  labs(title = "Science faceted by student grade level", x = "Exam Level", y = "Number of People who took the exam")
plot(graph2)

#Math
graph3 <- stud_perf_quantGPA %>%
  select(year, student_id_scram,grade_level,math_test_grade,math_perf_level) %>%
  group_by(grade_level, math_test_grade) %>%
  count() %>% 
  filter(!is.na(math_test_grade)) %>%
  ggplot(aes(x=math_test_grade, y = n)) +
  geom_point() +
  facet_wrap(~grade_level)  +
  labs(title = "Math faceted by student grade level", x = "Exam Level", y = "Number of People who took the exam")
plot(graph3)

#ELA
graph4 <- stud_perf_quantGPA %>%
  select(year, student_id_scram,grade_level,ela_test_grade,ela_perf_level) %>%
  group_by(grade_level, ela_test_grade) %>%
  count() %>% 
  filter(!is.na(ela_test_grade)) %>%
  ggplot(aes(x=ela_test_grade, y = n)) +
  geom_point() +
  facet_wrap(~grade_level) +
  labs(title = "Social Studies faceted by student grade level", x = "Exam Level", y = "Number of People who took the exam")
plot(graph4)

```
In any given grade level, we have a set of possible scores you could have: 
-not eligible
-eligible present
-eligible missing
```{r}
```
