---
title: "20170718-Map ProgramCodes to Schools"
author: "Anandini, Thoa"
date: "7/18/2017"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
#setwd('/data/nycdoe/')
```
This notebook contains code to fetch and join 2 types of data:

* 1) A table that maps program (codes and names) to school dbn's. Other details include: program description (short paragraph), program's interest, admission method, scores (eng, math, soc, sci, stm, ela), number of seats and number of applications per grade 9 and per grade 10, and other requirements. 

* 2) A "High School Directory" table, which helps map school dbn's to school names. Other details include: info about the location and building, contact info & website, grade info, public transport to get there, campus name, total number of students, school type/overview/program highlights, partner info.

## Match program codes to school codes

#### Read and clean up data:

We read in the data that map programs (name & code) to schools (name & code).

* Note 1: from OpenData NYC, we are able to obtain data of three years: 2016, 2014-15, 2013-14. 

* Note 2: we read in two types of tables. One maps program codes to dbn's. The other maps dbn's to school names.

```{r}
library(readr)
codes2016 <- read_csv(url('https://data.cityofnewyork.us/api/views/ge8j-uqbf/rows.csv?accessType=DOWNLOAD'))
codes201415 <- read_csv(url('https://data.cityofnewyork.us/api/views/by6m-6zpb/rows.csv?accessType=DOWNLOAD'))
codes201314 <- read_csv(url('https://data.cityofnewyork.us/api/views/i9pf-sj7c/rows.csv?accessType=DOWNLOAD'))

names2016 <- read_csv(url('https://data.cityofnewyork.us/api/views/7crd-d9xh/rows.csv?accessType=DOWNLOAD'))
names201415 <- read_csv(url('https://data.cityofnewyork.us/api/views/n3p6-zve2/rows.csv?accessType=DOWNLOAD'))
names201314 <- read_csv(url('https://data.cityofnewyork.us/api/views/u553-m549/rows.csv?accessType=DOWNLOAD'))

codes2016 %>% filter(!(dbn %in% names201314$dbn))
codes201415 %>% filter(!(dbn %in% names201415$dbn))  #identify the "insane" row and drop it.
codes201415 <-
  codes201415 %>%
  filter(dbn %in% names201415$dbn)
codes201314 %>% filter(!(DBN %in% names201314$DBN))
```

### Then we select variables that we need: 

And join and bind the tables across years.
```{r}
(schoolcodes2016 <-
  codes2016 %>%
  mutate(year = 2016) %>%
  full_join(names2016, by="dbn") %>%
  select(year, pgcode = code, pgname = program, dbn, school_name, interest, method)) %>%
  print(width = Inf)
(schoolcodes201415 <-
  codes201415 %>%
  mutate(year = 201415) %>%
  full_join(names201415, by="dbn") %>%
  select(year, pgcode = code, pgname = program, dbn, school_name, interest, method)) %>%
  print(width = Inf)
(schoolcodes201314 <-
  codes201314 %>%
  mutate(year = 201314) %>%
  full_join(names201314, by="DBN") %>%
  select(year,
         pgcode = CODE,
         pgname = PROGRAM,
         dbn = DBN,
         school_name = Printed_Name,
         interest = INTEREST,
         method = METHOD)) %>%
  print(width = Inf)

allschoolcodes <-
  rbind(schoolcodes201314,
        schoolcodes201415,
        schoolcodes2016) 

####### OLD CODE ##########################################
# allschoolcodes %>%
#   group_by(pgcode, dbn, year) %>%
#   count() %>%
#   filter(n != 1)
# allschoolcodes %>%
#   filter(dbn == "03M485")
# # there are 8 programs that have no code, and they all belong to year 2013-14, and they all belong to one of the 8 specialized school (already excluding LaGuardia M485)
# allschoolcodes %>% 
#   filter(is.na(pgcode))
# # we now try looking up these dbn's in the entire table, to see if they are present in other years or not.
# allschoolcodes %>%
#   filter(dbn %in% c("10X445",
#                     "31R605",
#                     "02M475",
#                     "05M692",
#                     "10X696",
#                     "13K430",
#                     "14K449",
#                     "28Q687") & year != 201314)
##############################################################
```

### Do some sanity check after the join and the bind:

```{r sanity check}
# this code helps us find out that programs offered at the 9 specialized schools do not have codes.
allschoolcodes %>% 
  filter(is.na(pgcode) | pgcode == "NO CODE")

# because there are no program codes there, we will drop these schools from our table.
# MANUAL WAY:
# allschoolcodes <-
#   allschoolcodes %>%
#   filter(!(dbn %in% c("03M485",
#                       "10X445",
#                       "31R605",
#                       "02M475",
#                       "05M692",
#                       "10X696",
#                       "13K430",
#                       "14K449",
#                       "28Q687")))
# SMARTER WAY:
allschoolcodes <-
  allschoolcodes %>%
  filter(!is.na(pgcode) & pgcode != "NO CODE")

# now every program appears only once per school in each year.
allschoolcodes %>%
  group_by(pgcode, dbn, year) %>%
  count() %>%
  filter(n != 1)

# even more precise: each program is unique to each school
allschoolcodes %>%
  group_by(pgcode) %>%
  summarize(num_schools = length(unique(dbn))) %>%
  filter(num_schools != 1)

# but one school may have multiple programs
allschoolcodes %>%
  group_by(dbn) %>%
  summarize(num_programs = length(unique(pgcode))) %>%
  filter(num_programs != 1)
```

### Ready!

The following code prepares a final-version of the program code table. We also include a sample code for joining our program code table to any tables of your need.

```{r map programs to schools in app_rate tables}
# because each program is unique to each school, we now collapse the table so that there's only 1 row per programcode
minimal_allschoolcodes <-
  allschoolcodes %>%
  mutate(year = ifelse(year == 2016,
                       yes = 201600,
                       no = year)) %>%
  group_by(pgcode) %>%
  filter(year == max(year)) %>%
  mutate(year = ifelse(year == 201600,
                       yes = 2016,
                       no = year))
  

### Sample code for joining our table to any table of your need
# your_table <-
#   your_table %>%
#   left_join(minimal_allschoolcodes, by=c("your_named_program" = "pgcode")) %>%
#   select(year = year,
#          pgcode = your_named_program,
#          pgname,
#          dbn,
#          school_name,
#          etc.)
```


```{r play-around, include=FALSE}
# # # some programs might not be present through all 3 years.
# allschoolcodes %>%
#   group_by(pgcode, dbn) %>%
#   count() %>%
#   filter(n < 3) %>%
#   arrange(n)
# 
# # let's check what year(s) these programs miss.
# allschoolcodes %>%
#   select(pgcode, dbn) %>%
#   spread
# 
# allschoolcodes %>%
#   # group_by(pgcode, dbn) %>%
#   # summarize(total_year_vals = sum(year)) %>%
#   # filter(total_year_vals != 404745) %>%
#   # left_join(allschoolcodes, by=c("pgcode", "dbn")) %>%
#   mutate(missed_year = ifelse(year == 201314,
#                               yes = ifelse(year == 201415,
#                                            yes = ifelse(year == 2016,
#                                                         yes = NA,
#                                                         no = 2016),
#                                            no = 201415),
#                               no = 201314)) %>%
#   select(pgcode, dbn, year, missed_year) %>%
#   group_by(pgcode, dbn) %>%
#   summarize()
#  # our guess is 2016 has the most exhaustive list of program codes. Here's our check:
# 
#   mutate(what_year_missed = as.character(404745 - total_year_vals)) %>%
#   mutate(what_year_missed = ifelse(what_year_missed == "402729", 
#                                    yes="201314 and 201516", 
#                                    no= ifelse(what_year_missed == "20330",
#                                               yes="201314 and 2016")) %>%
#   select(-total_year_vals) %>%
#   filter(what_year_missed != 0)
```

