---
title: "Figures and Stuff"
author: "Keri and David"
date: "7/31/2017"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, root.dir = '/data/nycdoe/')
options(tibble.width=Inf)
library(ggplot2)
library(readr)
library(tidyverse)
library(reshape2)
library(stringr)
library(scales)
theme_set(theme_bw())

load('/data/nycdoe/clean_data/dont_overrite_me.Rdata')
```
Loading ethnicity data frame
```{r}
read_june_with_year <- function(filename) {
  year <- as.numeric(gsub('-.*$', '', basename(filename)))
  df <- read_csv(filename, col_types = cols(student_id_scram = col_character(), grade_level = col_integer(), ethnicity = col_character()))
  df <- select(df, student_id_scram, ethnicity)
}

files <- Sys.glob('/data/nycdoe/June Biog/*.csv')
ethnicity<- map_df(files, read_june_with_year)
```

Dataframes for all students 

```{r}
student_unique_performance <-
  avg_all_percentile %>%
  group_by(student_id_scram, grade_level) %>%
  summarize(percentile = max(performance))

#Cleaning up data
student_unique_performance_long <-
  student_unique_performance %>%
  filter(!(is.na(grade_level) | is.na(student_id_scram) | is.na(percentile))) %>%
  filter(grade_level >= 3)
 
student_unique_performance_wide <- 
  student_unique_performance_long %>%
  spread(grade_level, percentile)

colnames(student_unique_performance_wide) <- 
  c("student_id_scram","grade3","grade4","grade5","grade6",
    "grade7","grade8","grade9","grade10","grade11","grade12")
```

Plotting the correlation coefficients
Only works with True Survivors

```{r}
load('/data/nycdoe/clean_data/true_surv_corr.Rdata')
plot(performance_correlations_of_true_survivors)
```

FUNCTION FOR PLOTTING GRADES AGAINST EACH OTHER

```{r}
# use standard eval hack to plot variable columns
# more here: http://dplyr.tidyverse.org/articles/programming.html
plotit <- function(df, grade_from, grade_to) {
  grade_from <- enquo(grade_from)
  grade_to <- enquo(grade_to)
  
  df %>%
  select(student_id_scram,!!grade_from,!!grade_to)%>%
  filter(!(is.na(!!grade_from)|is.na(!!grade_to)))%>%
  mutate(bin=round((!!grade_from)/5)*5) %>%
  group_by(bin) %>%
  summarize(low25=quantile(!!grade_to, 0.25), mid = quantile(!!grade_to, 0.5), high75=quantile(!!grade_to,0.75), 
            low10=quantile(!!grade_to,0.10), high90=quantile(!!grade_to,0.9)) %>%
    
  ggplot(aes(x=bin, y=mid)) +
  geom_line()+
  geom_ribbon(aes(ymin=low25, ymax=high75), alpha=0.2, fill="blue") +
  geom_ribbon(aes(ymin=low10, ymax=high90), alpha= 0.1,fill="blue") +
  geom_abline(linetype = "dashed")
}
p <- plotit(student_unique_performance_wide, grade4, grade5)
p + xlab("4th Grade Performance") + ylab("5th Grade Performance") + ggtitle("Performance in 4th Grade against Performance in 5th Grade")
fourvsfive<- p

q <- plotit(student_unique_performance_wide, grade4, grade8)
q + xlab("4th Grade Performance") + ylab("8th Grade Performance") + ggtitle("Performance in 4th Grade against Performance in 8th Grade")
fourvseight <-q

r <- plotit(student_unique_performance_wide, grade4, grade11)
r + xlab("4th Grade Performance") + ylab("11th Grade Performance") + ggtitle("Performance in 4th Grade against Performance in 11th Grade")
fourvseleven <-r

p <- plotit(student_unique_performance_wide, grade4, grade12)
p + xlab("4th Grade Performance") + ylab("12th Grade Performance") + ggtitle("Performance in 4th Grade against Performance in 12th Grade")
fourvseleven <- p

```

Progression of top 10 percent students in 3rd grade through 12th grade

```{r}
top_10_in_3 <-
  student_unique_performance_wide %>%
  select(student_id_scram,grade3) %>%
  filter(!is.na(grade3))%>%
  mutate(bin=round(grade3/5)*5) %>%
  filter(bin >= 90) %>%
  select(student_id_scram)

wide_10_in_3 <- left_join(top_10_in_3, student_unique_performance_wide, by="student_id_scram")

long_10_in_3 <-
  wide_10_in_3 %>% 
  gather(key="grade_level", value="percentile", 2:11) %>%
  select(student_id_scram, grade_level, percentile) %>%
  filter(!is.na(percentile))%>%
  group_by(grade_level) %>% 
  summarize(lower=quantile(percentile, 0.25), mid = quantile(percentile, 0.5), upper=quantile(percentile,0.75), 
            low=quantile(percentile,0.10), high=quantile(percentile,0.9), count=n()) %>% 
  mutate(grade_level = as.numeric(substr(grade_level,6,7)))

top_10_from_3rd  <-
  long_10_in_3 %>%
  ggplot(aes(x = grade_level, y= mid)) +
  geom_line()+
  geom_point(aes(size=count))+
  geom_ribbon(aes(ymin=lower, ymax=upper), alpha=0.2, fill="blue") +
  geom_ribbon(aes(ymin=low, ymax=high), alpha= 0.1,fill="blue") +
  labs(title = "Progression of the Top 10 percent of 3rd graders until 12th grade", 
       x = "Grade Level", y = "Student Percentile") +
  scale_x_discrete(limits = c('','','3','4','5','6','7','8','9','10','11','12'))
  
plot(top_10_from_3rd)

```

Progression of bottom 10 percent students in 3rd grade through 12th grade

```{r}
bot_10_in_3 <-
  student_unique_performance_wide %>%
  select(student_id_scram,grade3) %>%
  filter(!is.na(grade3))%>%
  mutate(bin=round(grade3/5)*5) %>%
  filter(bin <= 10) %>%
  select(student_id_scram)

wide_10_in_3 <- left_join(bot_10_in_3, student_unique_performance_wide, by="student_id_scram")

long_10_in_3 <-
  wide_10_in_3 %>% 
  gather(key="grade_level", value="percentile", 2:11) %>%
  select(student_id_scram, grade_level, percentile) %>%
  filter(!is.na(percentile))%>%
  group_by(grade_level) %>% 
  summarize(lower=quantile(percentile, 0.25), mid = quantile(percentile, 0.5), upper=quantile(percentile,0.75), 
            low=quantile(percentile,0.10), high=quantile(percentile,0.9), count=n()) %>% 
  mutate(grade_level = as.numeric(substr(grade_level,6,7)))

bot_10_from_3rd  <-
  long_10_in_3 %>%
  ggplot(aes(x = grade_level, y= mid)) +
  geom_line()+
  geom_point(aes(size=count))+
  geom_ribbon(aes(ymin=lower, ymax=upper), alpha=0.2, fill="red") +
  geom_ribbon(aes(ymin=low, ymax=high), alpha= 0.1,fill="red") +
  labs(title = "Progression of the Bottom 10 percent of 3rd graders until 12th grade", 
       x = "Grade Level", y = "Student Percentile") +
  scale_x_discrete(limits = c('','','3','4','5','6','7','8','9','10','11','12'))
  
plot(bot_10_from_3rd)
```
Bottom 10 Percentile Progression 
Sampled equally
```{r}
#Selecting student IDs of bottom 10 percent students
bot_10_in_3 <-
  student_unique_performance_wide %>%
  select(student_id_scram,grade3) %>%
  filter(!is.na(grade3))%>%
  mutate(bin=round(grade3/5)*5) %>%
  filter(bin <= 10) %>%
  select(student_id_scram)

#Combining their performance and their ethnicity
bot_wide_10_in_3 <- left_join(bot_10_in_3, student_unique_performance_wide, by="student_id_scram")
eth_bot_10 <- left_join(bot_wide_10_in_3, ethnicity, by="student_id_scram")

#Selecting student IDs of students who only has one ethnicity
one_eth_bot_10_id <-
  eth_bot_10 %>%
  group_by(student_id_scram, ethnicity) %>%
  unique() %>%
  ungroup() %>%
  group_by(student_id_scram) %>%
  summarize(count=n()) %>% 
  filter(count == 1) %>%
  select(student_id_scram)

one_eth_bot_10 <- left_join(one_eth_bot_10_id, eth_bot_10, by="student_id_scram")

#Sampling equal numbers for each student
set.seed(11)
sample_eth_bot_10_id <- 
  one_eth_bot_10 %>%
  filter(ethnicity != '.', ethnicity != 'Native American', ethnicity != 'Multi-Racial') %>%
  select(student_id_scram, ethnicity) %>% 
  group_by(student_id_scram) %>%
  unique() %>%
  group_by(ethnicity) %>%
  unique() %>% 
  sample_n(1000) %>%
  ungroup() %>%
  select(student_id_scram)

sample_eth_bot_10 <- left_join(sample_eth_bot_10_id, eth_bot_10, by="student_id_scram")

sample_eth_bot_10 %>% group_by(student_id_scram) %>%
  unique() %>%
  group_by(ethnicity) %>%
  count() %>%  View

long_eth_bot_10 <-
  sample_eth_bot_10 %>% 
  gather(key="grade_level", value="percentile", 2:11) %>%
  filter(!(is.na(percentile)))%>%
  filter(!(ethnicity=='.')) %>% 
  select(student_id_scram, grade_level, percentile, ethnicity) %>%
  mutate(grade_level = as.numeric(substr(grade_level,6,7))) %>%
  group_by(grade_level, ethnicity) %>%
  summarize(lower=quantile(percentile, 0.25), mid = quantile(percentile, 0.5), upper=quantile(percentile,0.75), 
            low=quantile(percentile,0.10), high=quantile(percentile,0.9),count=n())

long_eth_bot_10 %>%
  ggplot(aes(x = grade_level, y= mid)) +
  geom_line()+
  geom_ribbon(aes(ymin=lower, ymax=upper), alpha=0.2, fill="red") +
  geom_ribbon(aes(ymin=low, ymax=high), alpha= 0.1,fill="red") +
  geom_point(aes(size=count))+
  facet_wrap(~ethnicity) +
  labs(title = "Progression of the Bottom 10 percent of 3rd graders until 12th grade\nfaceted by ethnicity", 
       x = "Grade Level", y = "Student Percentile")
 
```
Top 10 Percentile Progression 
Sampled equally
```{r}
#Selecting student IDs of toptom 10 percent students
top_10_in_3 <-
  student_unique_performance_wide %>%
  select(student_id_scram,grade3) %>%
  filter(!is.na(grade3))%>%
  mutate(bin=round(grade3/5)*5) %>%
  filter(bin >= 90) %>%
  select(student_id_scram)

#Combining their performance and their ethnicity
top_wide_10_in_3 <- left_join(top_10_in_3, student_unique_performance_wide, by="student_id_scram")
eth_top_10 <- left_join(top_wide_10_in_3, ethnicity, by="student_id_scram")

#Selecting student IDs of students who only has one ethnicity
one_eth_top_10_id <-
  eth_top_10 %>%
  group_by(student_id_scram, ethnicity) %>%
  unique() %>%
  ungroup() %>%
  group_by(student_id_scram) %>%
  summarize(count=n()) %>% 
  filter(count == 1) %>%
  select(student_id_scram)

one_eth_top_10 <- left_join(one_eth_top_10_id, eth_top_10, by="student_id_scram")

#Sampling equal numbers for each student
set.seed(11)
sample_eth_top_10_id <- 
  one_eth_top_10 %>%
  filter(ethnicity != '.', ethnicity != 'Native American', ethnicity != 'Multi-Racial') %>%
  select(student_id_scram, ethnicity) %>% 
  group_by(student_id_scram) %>%
  unique() %>%
  group_by(ethnicity) %>%
  unique() %>% 
  sample_n(1000) %>%
  ungroup() %>%
  select(student_id_scram)

sample_eth_top_10 <- left_join(sample_eth_top_10_id, eth_top_10, by="student_id_scram")

sample_eth_top_10 %>% group_by(student_id_scram) %>%
  unique() %>%
  group_by(ethnicity) %>%
  count() %>%  View

long_eth_top_10 <-
  sample_eth_top_10 %>% 
  gather(key="grade_level", value="percentile", 2:11) %>%
  filter(!(is.na(percentile)))%>%
  filter(!(ethnicity=='.')) %>% 
  select(student_id_scram, grade_level, percentile, ethnicity) %>%
  mutate(grade_level = as.numeric(substr(grade_level,6,7))) %>%
  group_by(grade_level, ethnicity) %>%
  summarize(lower=quantile(percentile, 0.25), mid = quantile(percentile, 0.5), upper=quantile(percentile,0.75), 
            low=quantile(percentile,0.10), high=quantile(percentile,0.9),count=n())

long_eth_top_10 %>%
  ggplot(aes(x = grade_level, y= mid)) +
  geom_line()+
  geom_ribbon(aes(ymin=lower, ymax=upper), alpha=0.2, fill="blue") +
  geom_ribbon(aes(ymin=low, ymax=high), alpha= 0.1,fill="blue") +
  geom_point(aes(size=count))+
  facet_wrap(~ethnicity) +
  labs(title = "Progression of the Top 10 percent of 3rd graders until 12th grade\nfaceted by ethnicity", 
       x = "Grade Level", y = "Student Percentile")
 
```
