---
title: "Untitled"
author: "Keri and David"
date: "7/25/2017"
output: html_document
---
Necessary Set-up Stuff
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, root.dir = '/data/nycdoe/')
options(tibble.width=Inf)
library(ggplot2)
library(readr)
library(tidyverse)
library(reshape2)
library(stringr)
```
Loading in student year, student_id, and grade level from June Biog
```{r}
read_june_with_year <- function(filename) {
  year <- as.numeric(gsub('-.*$', '', basename(filename)))
  df <- read_csv(filename, col_types = cols(student_id_scram = col_character(), grade_level = col_integer()))
  df$year <- year
  df <- select(df, year, student_id_scram, grade_level)
}
files <- Sys.glob('/data/nycdoe/June Biog/*.csv')
level <- map_df(files, read_june_with_year)
```
Loading in Math and ELA stuff
```{r}
read_ela_math_with_year <- function(filename) {
  year <- as.numeric(gsub('-.*$', '', basename(filename)))
  df <- read_csv(filename, col_types = cols(student_id_scram = col_character()))
  df$year <- year
  df <- select(df, year, student_id_scram, ela_scale_score, math_scale_score)
}
files <- Sys.glob('/data/nycdoe/Math and ELA/*.csv')
ela_math <- map_df(files, read_ela_math_with_year)
```
Loading Regents Data
```{r}
load('/data/nycdoe/clean_data/highscool_performance.Rdata')
high_scores <- full_join(level, average_percentile, by = c("year","student_id_scram"))
```
Combine dataframes
```{r}
elem_scores <- full_join(level, ela_math, by = c("year","student_id_scram"))
```
Filtering the Dataframes
```{r}
#Adding the percent rank for each student per year
elem_scores_v2 <-
  elem_scores %>% 
  filter(!(is.na(ela_scale_score)&is.na(math_scale_score))) %>%
  mutate(ela_percent = (percent_rank(ela_scale_score))*100) %>%
  mutate(math_percent = (percent_rank(math_scale_score))*100)

#Gathering the dataset for dataframe for average to easily be computed
elem_scores_v3 <-
  elem_scores_v2 %>% 
  group_by(year,student_id_scram,grade_level) %>%
  select(year, student_id_scram, grade_level, ela_percent, math_percent) %>%
  gather(subject, percentile, ela_percent:math_percent, factor_key=TRUE)

#Average the percentiles for each student per year
final_elem_scores <-
  elem_scores_v3 %>%
  group_by(year,student_id_scram,grade_level) %>%
  filter(!is.na(percentile)) %>%
  summarize(average_percentile=mean(percentile))
```

Picking students who transitioned from 3rd to 4th grade and 
plotting their percentile change
```{r}
#grab students with grade level 3 and 4 ONLY
intermediate <- 
  final_elem_scores %>% 
  ungroup() %>%
  filter(grade_level==4 | grade_level ==5) %>%
  select(-year)

#grab student ids who appears twice with grade level 3 and 4
#not necessarily filter our students who repeated the year
id <- 
  intermediate %>% 
  group_by(student_id_scram) %>% 
  summarize(count = n()) %>% 
  filter(count == 2) %>% 
  ungroup %>% 
  select(student_id_scram)

#combined intermediate dataframe, still needs to filter out students who repeated same year
intermediate2 <- left_join(id, intermediate, by="student_id_scram")

#grab student ids who only appeared once each grade level (non-repeats)
id2 <- intermediate2 %>% 
  group_by(student_id_scram, grade_level) %>% 
  count() %>% 
  filter(n==1) %>% 
  ungroup() %>% 
  select(student_id_scram, grade_level)

#combine the student ids with the prefiltered dataframe
elem_scores_v5 <- left_join(id2, intermediate2, by=c("student_id_scram","grade_level"))

#spread the data frame on grade level
elem_scores_v6 <-
  elem_scores_v5 %>%
  spread(key = grade_level, value = average_percentile) 
colnames(elem_scores_v6) = c("student_id_scram","three_grade","four_grade")

elem_scores_v6 %>%
  ggplot(aes(x=(three_grade), y=(four_grade))) +
  geom_point(alpha = 0.01) 

```

Picking students who transition from 9th to 10th grade and 
plotting their percentile change
```{r}
#grab students with grade level 3 and 4 ONLY
inter_high <- 
  high_scores %>%
  filter(grade_level==9 | grade_level ==10) %>%
  select(-year)

#grab student ids who appears twice with grade level 3 and 4
#not necessarily filter our students who repeated the year
id_high <- 
  inter_high %>% 
  group_by(student_id_scram) %>% 
  summarize(count = n()) %>% 
  filter(count == 2) %>% 
  ungroup %>% 
  select(student_id_scram)

#combined intermediate dataframe, still needs to filter out students who repeated same year
inter2_high <- left_join(id_high, inter_high, by="student_id_scram")

#grab student ids who only appeared once each grade level (non-repeats)
id2_high <- inter2_high %>% 
  group_by(student_id_scram, grade_level) %>% 
  count() %>% 
  filter(n==1) %>% 
  ungroup() %>% 
  select(student_id_scram, grade_level)

#combine the student ids with the prefiltered dataframe
high_scores2 <- left_join(id2_high, inter2_high, by=c("student_id_scram","grade_level"))

#spread the data frame on grade level
high_scores3 <-
  high_scores2 %>%
  spread(key = grade_level, value = average_percentile) 
colnames(high_scores3) = c("student_id_scram","nine_grade","ten_grade")

high_scores3 %>%
  ggplot(aes(x=(nine_grade), y=(ten_grade))) +
  geom_point(alpha=0.01) +
  geom_smooth()

```


```{r}
all_scores <-
  full_join(high_scores, final_elem_scores, by=c("year","student_id_scram","grade_level"))

tidy_all_percentile <-
  all_scores %>%
  group_by(student_id_scram, year, grade_level) %>%
  gather(subject, percentile, 4:5)

avg_all_percentile <- 
  tidy_all_percentile %>%
  filter(!is.na(percentile)) %>%
  group_by(student_id_scram, year, grade_level) %>%
  summarize(performance = mean(percentile))

#AVERAGE JOE : 100002032 #

save(avg_all_percentile, file = '/data/nycdoe/clean_data/avg_all_long.Rdata')
```

```{r}
avg_all_long <- avg_all_percentile

avg_all_wide <- 
  avg_all_percentile %>%
  group_by(year,student_id_scram) %>%
  spread(key = grade_level, value = performance)
```